---
title: "Glowacki et al. 2026"
format: html
editor: visual
execute:
  warning: false
  dir: "C:/Users/sfield3/OneDrive - University of Wyoming/FIELD LAB HOME/PROJECT/SERIATION_MULTIVERSE"
---

# Research compendium for Glowacki et al. (2026)

This R-based research compendium supports the analyses from Seriation Multiverse, which shows that.....

View the raw source code and output at: provide link

**When using the code included in this research compendium, please cite *all* of the following:**

> Glowacki, Donna M., Katherine Portman, and Sean Field. 2026. *Methods Matters: Assessing the Seriation Multiverse in the U.S. Southwest*. Journal of Archaeological Method and Theory, Under Review.
>
> > Glowacki, Donna M., Katherine Portman, and Sean Field. 2026. *Research Compendium for Methods Matters: Assessing the Seriation Multiverse in the U.S. Southwest*. Journal of Archaeological Method and Theory. \[DOI\]

## 1.0 Set Up Environment

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "C:/Users/sfield3/OneDrive - University of Wyoming/FIELD LAB HOME/PROJECT/SERIATION_MULTIVERSE")
```

```{r}
#| warning: false

packages <-c('terra','sf','ggplot2','tidyterra','dplyr','tidyr','stringr','gifski','gganimate','ggforce','ggpubr','scales','lmtest','tseries','FedData','tidyverse','compiler','msm','snowfall','parallel','doParallel','ggridges')

for(p in packages) if(p %in% rownames(installed.packages()) == F) { install.packages(p) }
for(p in packages) suppressPackageStartupMessages(library(p,quietly=T,character.only=T))

theme_set(theme_bw())
```

### 1.1 Unique Functions

`contingency_table` : creates a contingency table (pair-wise presence & absence) of site occupation for each year

```{r}
contingency_table <- function(data1, data2, row) {
  contingency <- matrix(c(0,0,0,0), nrow = 2,
                        dimnames = list(dim1 = c("present", "absent"),
                                        dim2 = c("present", "absent")))
  for (i in 2:ncol(data1)) {
    if (data1[row, i] == 1) {
      if (data2[row, i] == 1) {contingency[1,1] <- contingency[1,1] + 1}
      else {contingency[1,2] <- contingency[1,2] + 1}
    } else {
      if (data2[row, i] == 1) {contingency[2,1] <- contingency[2,1] + 1}
      else {contingency[2,2] <- contingency[2,2] + 1}
    }
  }
  return (contingency)
}
```

`BR` : Brainerd-Robinson formula, developed by Peeples, Matthew A. (2011) R Script for Calculating the Brainerd-Robinson Coefficient of Similarity and Assessing Sampling Error. Electronic document, http://www.mattpeeples.net/br.html, accessed May 26, 2026. Access documentation [here](https://www.mattpeeples.net/BR.html) and code [here](https://www.mattpeeples.net/BR.R).

```{r}
BR <- function(x) {
  rd <- dim(x)[1]
  results <- matrix(0,rd,rd)
  for (s1 in 1:rd) {
    for (s2 in 1:rd) {
      x1Temp <- as.numeric(x[s1, ])
      x2Temp <- as.numeric(x[s2, ])
      br.temp <- 0
      results[s1,s2] <- 200 - (sum(abs(x1Temp - x2Temp)))}}
  row.names(results) <- row.names(x)
  colnames(results) <- row.names(x)
return(results)}
```

`updf`: Uniform Probability Density Analysis, developed as an R function by Peeples (2021) based on the approach published by Ortman (2016). Repository for code can be found [here](https://github.com/mpeeples2008/UniformProbabilityDensityAnalysis). Find Ortman's (2016) original work [here](https://link.springer.com/article/10.1007/s10816-014-9227-6).

```{r}
updf <- function(site,cer.type,ct,start,end,chron,interval=5,cutoff=0.1,min.period=25) {
  
  ## Remove any type with a count of 0 from all vectors
  trim <- which(ct>0)
  cer.type <- as.vector(cer.type[trim])
  site <- as.vector(site[1])
  start <- start[trim]
  end <- end[trim]
  ct <- ct[trim]
  chron <- chron[trim]
  
  
  ## define function for rounding by base defined in "interval" argument and round start and end dates for each type
  mround <- function(x,base){base*round(x/base)}
  start <- mround(start,interval)
  end <- mround(end,interval)
  
  ## find minimal ceramic intervals based on rounded type date overlaps
  years <- unique(sort(c(start,end)))
  period <- do.call(rbind,foreach(i=1:(length(years)-1)) %dopar% rbind(c(years[i],years[i+1]))) 
  # Ensure that no period is shorter than the specificed min.period argument and combine periods to satisfy this requirement
  if (length(period)>2) {
    per.short <- foreach(i=1:(nrow(period)-1)) %dopar% (period[i+1,1]-period[i,1])
    comb.per <- which(per.short<min.period)
    if (length(comb.per)>0){ period <- period[-comb.per,]
    for (i in 1:(nrow(period)-1)) {period[i,2] <- period[i+1,1]}}
  }  
  
  ## Define lists of ceramic period lengths and site period lengths
  cer.per <- foreach(i=1:length(start)) %dopar% (seq((start[i]:end[i]))+(start[i]-1))
  per.len <- foreach(i=1:nrow(period)) %dopar% (seq(period[i,1]:period[i,2])+(period[i,1])-1)
  
  ## Create Uniform Distance dataframe via parallel function
  ucalc <- function(a,b) {
    out <- (length(intersect(a,b))-1)/(length(a)-1)
    if (out<0) {out <- 0}
    return(out)}
  
  udist <- foreach(a=cer.per,.combine='rbind') %:% foreach(b=per.len,.combine='c') %dopar% (ucalc(a,b))
  if (length(udist)==1) {udist <- as.matrix(udist)}
  colnames(udist) <- paste('AD',period[,1],'-',period[,2],sep='')
  
  ## Create dataframe of prior values and sum to calculate prior probabilities
  prior.tab <- udist*ct
  if (ncol(prior.tab)>1) {
    if (length(prior.tab[which(chron==1),])==ncol(prior.tab)) {prior <- prior.tab[which(chron==1),]/sum(prior.tab[which(chron==1),])} else
    {prior <- colSums(prior.tab[which(chron==1),])/sum(prior.tab[which(chron==1),])}} else {
      prior <- udist[1,]
    }
  prior[is.na(prior)] <- 0
  
  ## Create dataframe of count probabilities
  pij <- sweep(prior.tab,2,colSums(prior.tab),'/')
  pij[is.na(pij)] <- 0
  
  ## Create dataframe of probabilities based on uniform distribution
  uij <- sweep(udist,2,colSums(udist),'/')
  uij[is.na(uij)] <- 0
  
  ## Create dataframe of standard deviations of uniform distribution probabilities
  sd.t <- sqrt(sweep((uij*(1-uij)),2,colSums(ceiling(udist)),'/'))
  
  ## Create dataframe of conditionals and calculate conditional probabilities
  cij <- dnorm(pij,uij,sd.t)
  cij[which(uij-sd.t==1)] <- dnorm(1,1,1) # For intervals with only one ceramic type, set to standard normal
  is.na(cij) <- !is.finite(cij) 
  
  # calculate conditional probability and rescale from 0-1
  conditional <- apply(cij,2,mean,na.rm=T)
  conditional[is.na(conditional)] <- 0
  if (sum(conditional)>0) {conditional <- conditional/sum(conditional)}
  
  ## Calculate posterior proportions and remove any generated NA
  posterior <- foreach(i=1:length(prior),.combine='c') %dopar% ((prior[i]*conditional[i]))
  posterior <- posterior/sum(posterior)
  posterior[is.na(posterior)] <- 0
  
  ## Deal with edge cases where sum of posterior probabilities = 0 or conditional = prior
  if(sum(posterior)==0) {posterior <- prior}
  if (identical(conditional,prior)) {posterior <- prior}
  
  # calculated beginning (lwr) and ending (upr) dates based on the first and last period with posterior probability about the selected threshold
  lwr <- period[min(which(posterior>cutoff*max(posterior))),1]
  upr <- period[max(which(posterior>cutoff*max(posterior))),2]
  
    ## Create output list and return
  out.list <- list(site=site,prior=prior,posterior=posterior,conditional=conditional,period=period,samp.size=sum(ct),occ=cbind(lwr,upr))
  return(out.list)
}
```

`nsj_ca_seriation`: Correspondence analysis based seriation developed by Glowacki and Field (2023), which can be accessed [here](https://link.springer.com/article/10.1007/s10816-014-9227-6).

```{r}
nsj_ca_seriation <- function(test_ceramics){
  cal_allprop <- read.csv("https://raw.githubusercontent.com/sfield2/NSJ-MV-CeramicSeriation/refs/heads/main/DATA/VEP_calibration_proportions.csv",header=T,fileEncoding = 'UTF-8-BOM')%>%
    as.matrix()
  
  #### Merge calibration and sample data
  test <-rbind(cal_allprop,test_ceramics)

  #### Run BR and extract relevant data
  test_br_out <- as.data.frame(BR(test))
  test_br_out <- test_br_out[11:nrow(test),1:10]

  ### normalize
  seriation_occupation<- as.data.frame(matrix(NA,nrow(test_ceramics),10))
  colnames(seriation_occupation) <- 1:10
  rownames(seriation_occupation) <- rownames(test_ceramics)
  
  for(i in 1:nrow(seriation_occupation)){
    for(j in 1:ncol(seriation_occupation)){
      seriation_occupation[i,j] <- (test_br_out[i,j]-rowMeans(test_br_out[i,]))/sd(test_br_out[i,])}
  }
  
  rm(cal_allprop)
return(seriation_occupation)
}
```

`rebin`: This function takes an annual occupational time series derived from a seriation method and re-attributes the occupations to date ranges (phase assignments) of a different seriation method.

```{r}
rebin <- function(seriation_annual, ranges){

  # Split ranges into numeric start/end
  parsed <- ranges %>%
    separate(ranges, into = c("start", "end"), sep = "-") %>%
    mutate(
      start = as.numeric(start),
      end   = as.numeric(end)
    )
  
  seriation_annual_rebinned <- seriation_annual %>%
    group_by(Site) %>%
    # For each period, check if any occupation exists in that range
    mutate(bins = sapply(Year, function(y) {
            active_periods <- parsed %>%
              mutate(in_range = y >= start & y <= end) %>%
              filter(in_range)
            
            if (nrow(active_periods) == 0) return(NA)
            has_occ <- any(Occupation > 0 & 
                         Year >= active_periods$start & 
                         Year <= active_periods$end)
            if (has_occ) {return(1)} 
              else {return(0)}
        })) %>%
    ungroup()
  return(seriation_annual_rebinned)
  }
```

`sort_ceramics` : This function summarizes the pottery counts for all `sites` and calculates the proportion of each type based on the total number in each ware.

```{r}
sort_ceramics <- function(all_pottery,sites){
  gray <- filter(all_pottery,Site.Number %in% sites$Site.Number)%>%
    select("Chapin.Gray","Moccasin.Gray","Mancos.Gray")%>%
    summarise(across(where(is.numeric),sum))%>%
    t()%>%
    as.data.frame()%>%
    rownames_to_column(var="Type")%>%
    setNames(c("Type","Count"))%>%
    mutate(Prop=Count/(sum(Count)),
         Ware="Gray Ware")

  corr <- filter(all_pottery,Site.Number %in% sites$Site.Number)%>%
    select("Mancos.Corrugated","Dolores.Corrugated","Mesa.Verde.Corrugated")%>%
    summarise(across(where(is.numeric),sum))%>%
    t()%>%
    as.data.frame()%>%
    rownames_to_column(var="Type")%>%
    setNames(c("Type","Count"))%>%
    mutate(Prop=Count/(sum(Count)),
         Ware="Corrugated")

  bw <- filter(all_pottery,Site.Number %in% sites$Site.Number)%>%
      select("Chapin.B.W","Piedra.B.W","Cortez.B.W","Mancos.B.W","McElmo.B.W","Mesa.Verde.B.W")%>%
    summarise(across(where(is.numeric),sum))%>%
    t()%>%
    as.data.frame()%>%
    rownames_to_column(var="Type")%>%
    setNames(c("Type","Count"))%>%
    mutate(Prop=Count/(sum(Count)),
         Ware="Painted B/W Ware")

  red <- filter(all_pottery,Site.Number %in% sites$Site.Number)%>%
    select("Abajo.R.O","Bluff.B.R","Deadmans.B.R")%>%
    summarise(across(where(is.numeric),sum))%>%
    t()%>%
    as.data.frame()%>%
    rownames_to_column(var="Type")%>%
    setNames(c("Type","Count"))%>%
    mutate(Prop=Count/(sum(Count)),
           Ware="Red Ware")

  pottery_plot<-rbind(gray,corr,bw,red)
  
  pottery_plot$Type <- factor(pottery_plot$Type,levels=c("Chapin.Gray","Chapin.B.W","Abajo.R.O","Piedra.B.W","Bluff.B.R","Moccasin.Gray","Mancos.Gray","Mancos.Corrugated","Cortez.B.W","Deadmans.B.R","Mancos.B.W","Dolores.Corrugated","McElmo.B.W","Mesa.Verde.Corrugated","Mesa.Verde.B.W"), labels=c("Chapin Gray","Chapin B/W","Abajo R/O","Piedra B/W","Bluff B/R","Moccasin Gray","Mancos Gray","Mancos\nCorrugated","Cortez B/W","Deadmans B/R","Mancos B/W","Dolores\nCorrugated","McElmo B/W","Mesa Verde\nCorrugated","Mesa Verde\nB/W"))
  
  pottery_plot[is.na(pottery_plot)] <- 0
  
  return(pottery_plot)
}
```

`cohen_kappa`: Calculates the inter-rater reliability for categorical variables based on a contingency table (pair-wise presence & absence) of site occupation for each year

```{r}
cohen_kappa <- function(data1,data2, row, n){
  contingency <- matrix(c(0,0,0,0), nrow = 2,
                        dimnames = list(dim1 = c("present", "absent"),
                                        dim2 = c("present", "absent")))
  for (i in 2:ncol(data1)) {
    if (data1[row, i] == 1) {
      if (data2[row, i] == 1) {contingency[1,1] <- contingency[1,1] + 1}
      else {contingency[1,2] <- contingency[1,2] + 1}
    } else {
      if (data2[row, i] == 1) {contingency[2,1] <- contingency[2,1] + 1}
      else {contingency[2,2] <- contingency[2,2] + 1}
    }
  }
  
  cm1 <- contingency[1,1]+contingency[2,1]

  cm2 <- contingency[1,2]+contingency[2,2]

  rm1 <- contingency[1,1]+contingency[1,2]

  rm2 <- contingency[2,1]+contingency[2,2]
  
  cohen_kappa_score <- (((cm1 * rm1)/(n)) + ((cm2 * rm2)/(n)))/n
  
  return(cohen_kappa_score)
}
```

`summarize_ceramics`:

```{r}
summarize_ceramics <- function(binned_data){
  for(i in 1:nrow(fifty_ranges)){
    begin <- fifty_ranges_summary[i,1]
    end <- fifty_ranges_summary[i,2]
    
    known_sites <- binned_data%>%
      filter(Year %in% begin:end,
         Occupation > 0)%>%
    distinct(Site)%>%
    setNames(c("Site.Number"))
    
    if(i ==1){
      ceramics <- sort_ceramics(master_pottery,known_sites)%>%
        mutate(Year_begin=begin,
               Year_end=end)
    }else{
      new_ceramics <- sort_ceramics(master_pottery,known_sites)%>%
        mutate(Year_begin=begin,
               Year_end=end)
      
      ceramics <- rbind(ceramics,new_ceramics)
    }
  }
  
  return(ceramics)
}
```

## 2.0 Import Study Area Extent, Sites, and Raw Ceramic Counts

1.  `sites`:

```{r}
#| echo: false
#| warning: false
sites <- st_read("./DATA/Spatial/ParkMesa_sites.gpkg")

# set number of sites for later analysis
site_threshold <- nrow(sites)

master_pottery <- read.csv("./DATA/Ceramics/PM_ceramics.csv")
```

```{r}

ca_ranges<- data.frame(ranges = c("0700-0799","0800-0879","0880-0919","0920-0979","0980-1019","1020-1059","1060-1099","1100-1179","1180-1249","1250-1280"), 
           period = 1:10)

pecosclass_ranges <- data.frame(ranges=c("0700-0899","0900-999","0900-1149","1050-1149","1150-1300"),period=c("P1","P1P2","P2","P2P3","P3"))

vep_ranges <-data.frame(ranges=c("0700-0724","0725-0799","0800-0839","0840-0879","0880-0919","0920-0979","0980-1019","1020-1059","1060-1099","1100-1139","1140-1179","1180-1224","1225-1259","1260-1280"),period=c("ph6","ph7","ph8","ph9","ph10","ph11","ph12","ph13","ph14","ph15","ph16","ph17","ph18","ph19"))

cybersw_ranges <- read.csv("./DATA/Ceramics/UPDA_Ceramic_TypesandDatesfromCyberSW.csv")%>%
  select(c("Type","BegDate","EndDate"))%>%
  setNames(c("Type","Begin","End"))

upda_ranges <-data.frame(ranges=c("0700-0799","0800-0874","0875-0924","0925-0974","0975-1049","1050-1174","1175-1200","1201-1250","1251-1280"),period=c("ph1","ph2","ph3","ph4","ph5","ph6","ph7","ph8","ph9"))

fifty_ranges <- data.frame(ranges=c("0700-0749","0750-0799","0800-0849","0850-0899","0900-0949","0950-0999","1000-1049","1050-1099","1100-1149","1150-1199","1200-1249","1250-1280"),period=c("ph1","ph2","ph3","ph4","ph5","ph6","ph7","ph8","ph9","ph10","ph11","ph12"))

fifty_ranges_summary <- data.frame(begin=c(700,750,800,850,900,950,1000,1050,1100,1150,1200,1250),end=c(749,799,849,899,949,999,1049,1099,1149,1199,1249,1280))
```

## 3.0 Read in Pre-analyzed Seriation Results

`vep`:

`pecosclass`:

```{r}
vep <- read.csv("./DATA/Occupations/vep_output.csv")
pecosclass <- read.csv("./DATA/Occupations/survey_output.csv")
```

## 4.0 Correspondence Analysis

Transform diagnostic sherds into proportions then run correspondence analysis.

```{r}
# if cut off on ceramics is >= 5
df_diag <- subset(master_pottery, select=c(Site.Number,Chapin.Gray,Moccasin.Gray,Mancos.Gray,Mancos.Corrugated,Dolores.Corrugated,Mesa.Verde.Corrugated,Chapin.B.W,Piedra.B.W,Cortez.B.W,Mancos.B.W,McElmo.B.W,Mesa.Verde.B.W,Abajo.R.O,Bluff.B.R,Deadmans.B.R))%>%
column_to_rownames("Site.Number")

df_diag_allprop <-prop.table(as.matrix(df_diag),1)*100

ca <- nsj_ca_seriation(df_diag_allprop)
```

## 5.0 UPDA

Spread master_pottery into types and assign counts

```{r}
diagnostic_list <- c("Abajo.R.O","Bluff.B.R","Chapin.B.W","Chapin.Gray","Cortez.B.W","Deadmans.B.R","Dolores.Corrugated","Mancos.B.W","Mancos.Corrugated","Mancos.Gray","McElmo.B.W","Mesa.Verde.B.W","Mesa.Verde.Corrugated","Moccasin.Gray","Piedra.B.W")

upda_ceramics <- master_pottery%>%
  select(-c(Source))%>%
  gather(Type, Count,-Site.Number)%>%
  left_join(cybersw_ranges,by="Type")%>%
  mutate(Chronology=ifelse(.$Type%in% diagnostic_list,1,0))%>%
  filter(Type!="Other",
         Count > 0)
```

```{r}
site_list <- unique(upda_ceramics$Site.Number)

## Run analysis for all sites
output_list <- list()

for (i in 1:length(site_list)) {
  site_sub <- upda_ceramics[which(upda_ceramics$Site == site_list[i]),]
  updf_out <- updf(site = site_sub$Site.Number,
                   cer.type=site_sub$Type,
                   ct=site_sub$Count,
                   start=site_sub$Begin,
                   end=site_sub$End,
                   chron=site_sub$Chronology,
                   interval=25,
                   cutoff=0.1,
                   min.period=10)
  output_list[[i]] <- updf_out
}

## Output estimated occupation ("occ" variable) for each site, read into csv
upda <- as.data.frame(matrix(nrow = length(site_list), ncol = 2, dimnames = list(site_list, c('Start', 'End'))))

for (i in 1:length(site_list)) {
  upda[i,1] <- output_list[[i]]$occ[1]
  upda[i,2] <- output_list[[i]]$occ[2]
}
```

## 6.0 Analysis

### 6.1 Transform Occupation assignments into Annual Reconstructions

```{r}
ca_annual <- ca%>%
  rownames_to_column(var = "Site.Number")%>%
  gather(key="Timeperiod", value = "Value",`1`:`10`)%>%
  mutate(range = ca_ranges$ranges[match(Timeperiod,ca_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise() %>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))%>%
  mutate(Occupation = ifelse(Occupation>0,1,0))%>%
  filter(Year>749,
         Year<1281)

pecosclass_annual_nontransition <- pecosclass%>%
  select(-c("P1P2","P2P3"))%>%
  gather(key="Timeperiod", value = "Value",-Site.Number)%>%
  mutate(range = pecosclass_ranges$ranges[match(Timeperiod,pecosclass_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise() %>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))

pecosclass_annual_transition <- pecosclass%>%
  select(c("Site.Number","P1P2","P2P3"))%>%
  gather(key="Timeperiod", value = "Value",-Site.Number)%>%
  mutate(range = pecosclass_ranges$ranges[match(Timeperiod,pecosclass_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise() %>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))

pecosclass_annual <- full_join(pecosclass_annual_nontransition,pecosclass_annual_transition, by = c("Site","Year"))%>%
  mutate(Occupation = rowSums(cbind(Occupation.x,Occupation.y),na.rm=T))%>%
  select(-c("Occupation.x","Occupation.y"))%>% 
  mutate(Occupation = replace_na(Occupation, 0))%>%
  filter(Year>749,
         Year<1281)

vep_annual <- vep%>%
  gather(key="Timeperiod", value = "Value",-Site.Number)%>%
  mutate(range = vep_ranges$ranges[match(Timeperiod,vep_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise()%>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))%>%
  mutate(Occupation = ifelse(Occupation>0,1,0))%>%
  filter(Year>749,
         Year<1281)


upda_annual <- upda%>%
  mutate(Value=1)%>%
  rownames_to_column(var = "Site.Number")%>%
  rowwise()%>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$Start,.$End,by=1)))%>%
  complete(Site,Year=750:1300,fill=list(Occupation=0))%>%
  filter(Year>749,
         Year<1281)


rm(pecosclass_annual_nontransition,pecosclass_annual_transition)
```

#### 6.1.1 Figure 2: Plot Summary Counts

```{r}
pecosclass_sum <- pecosclass_annual%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

ca_sum <- ca_annual%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

vep_sum <- vep_annual%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

upda_sum <- upda_annual%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))


# full_summary <- merge(pecosclass_sum,ca_sum,by="Year")%>%
#   merge(.,vep_sum,by="Year")%>%
#   merge(.,upda_sum,by="Year")%>%
#   setNames(c("Year","pecosclass","CA","VEP","UPDA"))%>%
#   gather(Method,Count,-Year)
# 
# full_summary$Method <- factor(full_summary$Method, levels=c("pecosclass","CA","UPDA","VEP"))

pecosclass_ranges_plot <- pecosclass_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))%>%
  mutate(phase = c("PI","PI/PII","PII","PII/PIII","PIII"))

pecosclass_ranges_plot1 <- pecosclass_ranges_plot[c(1,3,5),]
pecosclass_ranges_plot2 <- pecosclass_ranges_plot[c(2,4),]

ca_ranges_plot <- ca_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))%>%
  mutate(phase = c("ph1","ph2","ph3","ph4","ph5","ph6","ph7","ph8","ph9","ph10"))

upda_ranges_plot <- upda_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))

vep_ranges_plot <- vep_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))
  


p1<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=4,ymax=5),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1.25)+
    geom_text(data=pecosclass_ranges_plot1,aes(x = (start + end)/2,y = 4.5,label=phase,size=10))+
  geom_rect(data=pecosclass_ranges_plot2,aes(xmin=start,xmax=end,ymin=4.5,ymax=5),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1.25)+
  geom_text(data=pecosclass_ranges_plot2,aes(x = (start + end)/2,y = 4.75,label=phase,size=10))+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=2.75,ymax=3.75),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1.25)+
  geom_text(data=ca_ranges_plot,aes(x = (start + end)/2,y = 3.25,label=phase,size=10))+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1.5,ymax=2.5),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1.25)+
  geom_text(data=upda_ranges_plot,aes(x = (start + end)/2,y = 2,label=period,size=10))+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=0.25,ymax=1.25),col="#BFA373",fill="#F2CC8FA4",alpha=0.5,linewidth = 1.25)+
  geom_text(data=vep_ranges_plot,aes(x = (start + end)/2,y = 0.75,label=period,size=10))+
  annotate(geom="text",x=680,y=0.75,label="VEP")+
  annotate(geom="text",x=680,y=2,label="UPDA")+
  annotate(geom="text",x=680,y=3.25,label="CA")+
  annotate(geom="text",x=680,y=4.5,label="PC")+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,5),breaks=seq(0,5,by=1))+
  labs(x=NULL,
       y="Seriation Method\nPhases")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.1, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text.x=element_text(size=13),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = margin(5,5,0,2))

p2<-ggplot(data=pecosclass_sum,aes(Year,total_count))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#3d405b",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,80),breaks=seq(0,80,by=20))+
  labs(x=NULL,
       y="Site Count\nPecos Classification")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p3<-ggplot(data=ca_sum ,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#E07A5F",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,80),breaks=seq(0,80,by=20))+
  labs(x=NULL,
       y="Site Count\n(CA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p4<-ggplot(data=upda_sum ,aes(Year,total_count))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#81B29A",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,80),breaks=seq(0,80,by=20))+
  labs(x=NULL,
       y="Site Count\n(UPDA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p5<-ggplot(data=vep_sum ,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#BFA373",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,80),breaks=seq(0,80,by=20))+
  labs(x="Year",
       y="Site Count\n(VEP)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))


png('./FIGURES/Figure_2.png',height=16000,width=16000,res=1200)
ggarrange(p1,p2,p3,p4,p5, nrow=5,ncol=1,align="hv",heights=c(1.5,1,1,1,1))
dev.off()
```

### 6.3 Merge with spatial data

These data are used fro Figure 3.

```{r}
sites_stripped <- sites%>%
  select(c("NewSiteNum","geom"))%>%
  setNames(c("Site","geom"))

pecosclass1 <- spread(pecosclass_annual,Year,Occupation)
ca1 <- spread(ca_annual,Year,Occupation)
upda1 <- spread(upda_annual,Year,Occupation)
vep1 <- spread(vep_annual,Year,Occupation)

pecosclass_spatial <- merge(sites_stripped,pecosclass1,by="Site")
ca_spatial <- merge(sites_stripped,ca1,by="Site")
upda_spatial <- merge(sites_stripped,upda1,by="Site")
vep_spatial <- merge(sites_stripped,vep1,by="Site")

#st_write(pecosclass_spatial,"./OUTPUT/Spatial/pecosclass_spatial.gpkg")
#st_write(ca_spatial,"./OUTPUT/Spatial/ca_spatial.gpkg")
#st_write(upda_spatial,"./OUTPUT/Spatial/upda_spatial.gpkg")
#st_write(vep_spatial,"./OUTPUT/Spatial/vep_spatial.gpkg")
```

### 6.4 Disagreements

#### 6.4.1 Chi Squared

```{r}
#| warning: false 
row_names <- 750:1280

summary <- merge(ca_sum,vep_sum,by="Year")%>%
  merge(.,upda_sum,by="Year")%>%
  merge(.,pecosclass_sum,by="Year")%>%
  setNames(c("Year","CA","VEP","UPDA","pecosclass"))

chisq_p <- matrix(nrow = nrow(summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"],summary[i,c("VEP")],site_threshold-summary[i,c("VEP")],summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")]),nrow = 2),correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values.csv")
```

#### 6.4.2 Disagreement Frequencies

Two steps:

1.  Build into site-by-site time series
2.  Calculate pairwise disagreements

```{r}
#| warning: false 
pecosclass_t <- pecosclass_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)

vep_t <- vep_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)

ca_t <- ca_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)

upda_t <- upda_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)


row_names <- 750:1280

summary <- merge(ca_sum,vep_sum,by="Year")%>%
  merge(.,upda_sum,by="Year")%>%
  merge(.,pecosclass_sum,by="Year")%>%
  setNames(c("Year","CA","VEP","UPDA","pecosclass"))


col_names <- c("pecosclass_CA","pecosclass_UPDA","pecosclass_VEP","CA_UPDA", "CA_VEP", "UPDA_VEP")

percent_error <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(percent_error)) {
  #pecosclass vs. CA
  percent_error[i,c("pecosclass_CA")] <- (contingency_table(pecosclass_t, ca_t, i)[1,2] + 
                           contingency_table(pecosclass_t, ca_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. UPDA
  percent_error[i,c("pecosclass_UPDA")] <- (contingency_table(pecosclass_t, upda_t, i)[1,2] + 
                           contingency_table(pecosclass_t, upda_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. VEP
  percent_error[i,c("pecosclass_VEP")] <- (contingency_table(pecosclass_t, vep_t, i)[1,2] + 
                           contingency_table(pecosclass_t, vep_t, i)[2,1]) /site_threshold

  #CA vs. UPDA
  percent_error[i,c("CA_UPDA")] <- (contingency_table(ca_t, upda_t, i)[1,2] + 
                           contingency_table(ca_t, upda_t, i)[2,1]) /site_threshold
  
  #CA vs. VEP
  percent_error[i,c("CA_VEP")] <- (contingency_table(ca_t,vep_t, i)[1,2] + 
                           contingency_table(ca_t, vep_t, i)[2,1]) /site_threshold
  
  #UPDA vs. VEP
  percent_error[i,c("UPDA_VEP")] <- (contingency_table(upda_t, vep_t, i)[1,2] + 
                           contingency_table(upda_t, vep_t, i)[2,1]) /site_threshold
  
  
  
t <-chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  
  
}

write.csv(percent_error, "./OUTPUT/percent_error.csv",row.names=F)
```

#### 6.4.3 Table 3: Table of Disagreements

```{r}
percent_error <- read.csv("./OUTPUT/percent_error.csv")

disagreement_summary1<-percent_error%>%
  mutate(pecosclass_average_error = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         CA_average_error = (pecosclass_CA+CA_UPDA+CA_VEP)/3,
         UPDA_average_error = (pecosclass_UPDA+CA_UPDA+UPDA_VEP)/3,
         VEP_average_error = (pecosclass_VEP+CA_VEP+UPDA_VEP)/3,
         Year=750:1280)

table_3 <- t(sapply(disagreement_summary1, function(x) {
  c(mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE))
}))%>%
  as.data.frame()

all_summary <- disagreement_summary1%>%
  select(-c("Year"))%>%
  gather(method,value)%>%
  select(-c("method"))

table_3 <- rbind(table_3, "All_data"=list(mean=mean(all_summary$value),
          median=median(all_summary$value),
          sd=sd(all_summary$value),
          min=min(all_summary$value),
          max=max(all_summary$value)))
  

write.csv(table_3,"./OUTPUT/Table3_raw.csv")
```

#### 6.4.4 Figure 4: Plot Chi Squared and Disagreement Results

```{r}
#| warning: false 
chisq_p1 <- read.csv("./OUTPUT/chi_squared_p-values.csv")

disagreement_pecosclass1 <- disagreement_summary1%>%
  select(c("Year","pecosclass_average_error","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

disagreement_ca1 <- disagreement_summary1%>%
  select(c("Year","CA_average_error","pecosclass_CA","CA_UPDA","CA_VEP"))%>%
  gather(Method,error,-Year)

disagreement_upda1 <- disagreement_summary1%>%
  select(c("Year","UPDA_average_error","pecosclass_UPDA","CA_UPDA","UPDA_VEP"))%>%
  gather(Method,error,-Year)

disagreement_vep1 <- disagreement_summary1%>%
  select(c("Year","VEP_average_error","pecosclass_VEP","CA_VEP","UPDA_VEP"))%>%
  gather(Method,error,-Year)



### REFACTOR FOR CLEAN PLOTTING
disagreement_pecosclass1$Method <- factor(disagreement_pecosclass1$Method,levels=c("pecosclass_CA","pecosclass_UPDA","pecosclass_VEP","pecosclass_average_error"))

disagreement_ca1$Method <- factor(disagreement_ca1$Method,levels=c("pecosclass_CA","CA_UPDA","CA_VEP","CA_average_error"))

disagreement_upda1$Method <- factor(disagreement_upda1$Method,levels=c("pecosclass_UPDA","CA_UPDA","UPDA_VEP","UPDA_average_error"))

disagreement_vep1$Method <- factor(disagreement_vep1$Method,levels=c("pecosclass_VEP","CA_VEP","UPDA_VEP","VEP_average_error"))
###



p1<-ggplot(data=chisq_p1,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  labs(x=NULL,
       y="Chi-Square P-value",
       title="Average Disagreement by Method")+
  theme_minimal()+
  theme(#legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p2<-ggplot(data=disagreement_pecosclass1 ,aes(Year,error))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#6C6C8140","#6C6C8140","#6C6C8140","#3d405b"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y="Disagreement\n(Pecos Class)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p3<-ggplot(data=disagreement_ca1 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F40","#E07A5F40","#E07A5F40","#E07A5F"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y="Disagreement\n(CA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p4<-ggplot(data=disagreement_upda1 ,aes(Year,error))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#BCD1C580","#BCD1C580","#81B29A"))+
  labs(x=NULL,
       y="Disagreement\n(UPDA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p5<-ggplot(data=disagreement_vep1 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BFA37366","#BFA37366","#BFA37366","#BFA373"))+
  labs(x="Year",
       y="Disagreement\n(VEP)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))


png('./FIGURES/Figure_4.png',height=16000,width=12000,res=1200)
ggarrange(p1,p2,p3,p4,p5, nrow=5,ncol=1,align="hv",heights=c(0.75,1,1,1,1))
dev.off()
```

#### 6.4.5 Figure 5: Pair-wise Disagreements

```{r}
percent_error <- read.csv("./OUTPUT/percent_error.csv")%>%
  mutate(Year=750:1280)%>%
  gather(method,disagree,-Year)

pecosclass_ca <-filter(percent_error,method=="pecosclass_CA")
pecosclass_upda <-filter(percent_error,method=="pecosclass_UPDA")
pecosclass_vep <-filter(percent_error,method=="pecosclass_VEP")
ca_upda <-filter(percent_error,method=="CA_UPDA")
ca_vep <-filter(percent_error,method=="CA_VEP")
upda_vep <-filter(percent_error,method=="UPDA_VEP")


p1<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_ca,aes(Year,disagree),col="#8F5D5D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y="Disagreement",
       title ="Pecos Classs & CA")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title.y = element_text(size=15),
        axis.text.y=element_text(size=13),
        axis.text.x=element_blank(),
        plot.margin = margin(10,10,10,5))

p2<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_upda,aes(Year,disagree),col="#5F797B",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & UPDA")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p3<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#BFA373",fill="#F2CC8FA4",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_vep,aes(Year,disagree),col="#7E7267",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & VEP")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p4<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
    geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_upda,aes(Year,disagree),col="#B1967D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y="Disagreement",
       title ="CA & UPDA")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))


p5<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#BFA373",fill="#F2CC8FA4",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_vep,aes(Year,disagree),col="#D08F69",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="CA & VEP")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p6<-ggplot()+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#BFA373",fill="#F2CC8FA4",alpha=0.5,linewidth = 1)+
  geom_line(data=upda_vep,aes(Year,disagree),col="#A0AB87",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="UPDA & VEP")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

png('./FIGURES/Figure_5.png',height=14000,width=24000,res=1200)
ggarrange(p1,p2,p3,p4,p5,p6, nrow=2,ncol=3,align="hv")
dev.off()
```

### 6.5 Rebinning to Different Phase Configurations

#### 6.5.1 Re-bin to Pecos

```{r}
#| warning: false
ca_bintopecosclass <- rebin(ca_annual,pecosclass_ranges)
vep_bintopecosclass <- rebin(vep_annual,pecosclass_ranges)
upda_bintopecosclass <- rebin(upda_annual,pecosclass_ranges)
pecosclass_bintopecosclass <- pecosclass_annual

ca_bintopecosclass_sum <- ca_bintopecosclass%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

vep_bintopecosclass_sum <- vep_bintopecosclass%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

upda_bintopecosclass_sum <- upda_bintopecosclass%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

pecosclass_bintopecosclass_sum <- pecosclass_bintopecosclass%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

pecosclass_bins_summary <- merge(ca_bintopecosclass_sum,vep_bintopecosclass_sum,by="Year")%>%
  merge(.,upda_bintopecosclass_sum,by="Year")%>%
  merge(.,pecosclass_bintopecosclass_sum,by="Year")%>%
  setNames(c("Year","CA","VEP","UPDA","pecosclass"))%>%
  gather(Method,Count,-Year)
```

#### 6.5.2 Re-bin to VEP

```{r}
ca_bintovep <- rebin(ca_annual,vep_ranges)
pecosclass_bintovep <- rebin(pecosclass_annual,vep_ranges)
upda_bintovep <- rebin(upda_annual,vep_ranges)
vep_bintovep <- vep_annual

ca_bintovep_sum <- ca_bintovep%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

pecosclass_bintovep_sum <- pecosclass_bintovep%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

upda_bintovep_sum <- upda_bintovep%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

vep_bintovep_sum <- vep_bintovep%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

vep_bins_summary <- merge(ca_bintovep_sum,pecosclass_bintovep_sum,by="Year")%>%
  merge(.,upda_bintovep_sum,by="Year")%>%
  merge(.,vep_bintovep_sum,by="Year")%>%
  setNames(c("Year","CA","pecosclass","UPDA","VEP"))%>%
  gather(Method,Count,-Year)
```

#### 6.5.3 Re-bin to CA

```{r}
pecosclass_bintoca <- rebin(pecosclass_annual,ca_ranges)
vep_bintoca <- rebin(vep_annual,ca_ranges)
upda_bintoca <- rebin(upda_annual,ca_ranges)
ca_bintoca <- ca_annual

pecosclass_bintoca_sum <- pecosclass_bintoca%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

vep_bintoca_sum <- vep_bintoca%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

upda_bintoca_sum <- upda_bintoca%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

ca_bintoca_sum <- ca_bintoca%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

ca_bins_summary <- merge(pecosclass_bintoca_sum,vep_bintoca_sum,by="Year")%>%
  merge(.,upda_bintoca_sum,by="Year")%>%
  merge(.,ca_bintoca_sum,by="Year")%>%
  setNames(c("Year","pecosclass","VEP","UPDA","CA"))%>%
  gather(Method,Count,-Year)
```

#### 6.5.4 Re-bin to UPDA

```{r}
pecosclass_bintoupda <- rebin(pecosclass_annual,upda_ranges)
vep_bintoupda <- rebin(vep_annual,upda_ranges)
ca_bintoupda <- rebin(ca_annual,upda_ranges)
upda_bintoupda <- upda_annual

pecosclass_bintoupda_sum <- pecosclass_bintoupda%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

vep_bintoupda_sum <- vep_bintoupda%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

ca_bintoupda_sum <- ca_bintoupda%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

upda_bintoupda_sum <- upda_bintoupda%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

upda_bins_summary <- merge(pecosclass_bintoupda_sum,vep_bintoupda_sum,by="Year")%>%
  merge(.,ca_bintoupda_sum,by="Year")%>%
  merge(.,upda_bintoupda_sum,by="Year")%>%
  setNames(c("Year","pecosclass","VEP","UPDA","CA"))%>%
  gather(Method,Count,-Year)
```

#### 6.5.5 Figure 6: Plot Effects of Re-binning

```{r}
p1<-ggplot(data=pecosclass_bintopecosclass_sum,aes(Year,total_count))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#3d405b",lwd=1.5)+
  labs(title="Binned to Pecos Classification",
       x=NULL,
       y="Pecos Class Site Count")+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title.y = element_text(size=15),
        plot.margin = margin(5,5,0,2),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1))


p2<- ggplot(data=pecosclass_bintoca_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#3d405b",lwd=1.5)+
  labs(title="Binned to CA Phases",
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p3 <- ggplot(data=pecosclass_bintoupda_sum,aes(Year,total_count))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#3d405b",lwd=1.5)+
  labs(title="Binned to UPDA Phases",
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))


p4 <- ggplot(data=pecosclass_bintovep_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#3d405b",lwd=1.5)+
  labs(title="Binned to VEP Phases",
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p5 <- ggplot(data=ca_bintopecosclass_sum,aes(Year,total_count))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#e07a5f",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y="CA Site Counts")+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        plot.margin = margin(5,5,0,2))

p6 <- ggplot(data=ca_bintoca_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#e07a5f",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.margin = margin(5,5,0,2),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1))


p7 <- ggplot(data=ca_bintoupda_sum,aes(Year,total_count))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#e07a5f",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.margin = margin(5,5,0,2))


p8 <- ggplot(data=ca_bintovep_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#e07a5f",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.margin = margin(5,5,0,2))

p9 <- ggplot(data=upda_bintopecosclass_sum,aes(Year,total_count))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#81b29a",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y="UPDA Site Counts")+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        plot.margin = margin(5,5,0,2))

p10 <- ggplot(data=upda_bintoca_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#81b29a",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.margin = margin(5,5,0,2))

p11 <- ggplot(data=upda_bintoupda_sum,aes(Year,total_count))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#81b29a",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.margin = margin(5,5,0,2),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1))

p12 <- ggplot(data=upda_bintovep_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#81b29a",lwd=1.5)+
  labs(title=NULL,
       x=NULL,
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.margin = margin(5,5,0,2))

p13 <- ggplot(data=vep_bintopecosclass_sum,aes(Year,total_count))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#BFA373",lwd=1.5)+
  labs(title=NULL,
       x="Year",
       y="VEP Site Counts")+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title = element_text(size=15),
        plot.margin = margin(5,5,0,2))

p14 <- ggplot(data=vep_bintoca_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#BFA373",lwd=1.5)+
  labs(title=NULL,
       x="Year",
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.x = element_text(size=15),
        plot.margin = margin(5,5,0,2))

p15 <- ggplot(data=vep_bintoupda_sum,aes(Year,total_count))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#BFA373",lwd=1.5)+
  labs(title=NULL,
       x="Year",
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.x = element_text(size=15),
        plot.margin = margin(5,5,0,2))

p16 <- ggplot(data=vep_bintovep_sum,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#BFA373",lwd=1.5)+
  labs(title=NULL,
       x="Year",
       y=NULL)+
  scale_y_continuous(limits=c(0,80))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.x = element_text(size=15),
        plot.margin = margin(5,5,0,2),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1))

png('./FIGURES/Figure_6.png',height=12000,width=24000,res=1200)
ggarrange(p1,p2,p3,p4,
          p5,p6,p7,p8,
          p9,p10,p11,p12,
          p13,p14,p15,p16,
          nrow=4,ncol=4,align="hv")
dev.off()
```

### 6.6 Disagreeable Sites, Multicomponent Occupation, and Red Wares

1.  Find most disagreeable sites
2.  Find sites with the greatest number of components
3.  Look at ceramics of most disagreeable sites with high components in relationship to entire assemblage.
    1.  What happens to red wares?

```{r}
#| warning: false
### find out which sites most disagreeable
pecosclass_ca <- abs(pecosclass_t-ca_t)%>%
  select(-c("Year"))%>%
  t()%>%
  as.data.frame()%>%
  mutate(sums=rowSums(across(where(is.numeric)),na.rm=T))%>%
  select(c(sums))%>%
  rownames_to_column(var="Site.Number")%>%
  setNames(c("Site.Number","pecosclass_ca"))

pecosclass_upda <- abs(pecosclass_t-upda_t)%>%
  select(-c("Year"))%>%
  t()%>%
  as.data.frame()%>%
  mutate(sums=rowSums(across(where(is.numeric)),na.rm=T))%>%
  select(c(sums))%>%
  rownames_to_column(var="Site.Number")%>%
  setNames(c("Site.Number","pecosclass_upda"))

pecosclass_vep <- abs(pecosclass_t-vep_t)%>%
  select(-c("Year"))%>%
  t()%>%
  as.data.frame()%>%
  mutate(sums=rowSums(across(where(is.numeric)),na.rm=T))%>%
  select(c(sums))%>%
  rownames_to_column(var="Site.Number")%>%
  setNames(c("Site.Number","pecosclass_vep"))

ca_upda <- abs(ca_t-upda_t)%>%
  select(-c("Year"))%>%
  t()%>%
  as.data.frame()%>%
  mutate(sums=rowSums(across(where(is.numeric)),na.rm=T))%>%
  select(c(sums))%>%
  rownames_to_column(var="Site.Number")%>%
  setNames(c("Site.Number","ca_upda"))

ca_vep <- abs(ca_t-vep_t)%>%
  select(-c("Year"))%>%
  t()%>%
  as.data.frame()%>%
  mutate(sums=rowSums(across(where(is.numeric)),na.rm=T))%>%
  select(c(sums))%>%
  rownames_to_column(var="Site.Number")%>%
  setNames(c("Site.Number","ca_vep"))

upda_vep <- abs(upda_t-vep_t)%>%
  select(-c("Year"))%>%
  t()%>%
  as.data.frame()%>%
  mutate(sums=rowSums(across(where(is.numeric)),na.rm=T))%>%
  select(c(sums))%>%
  rownames_to_column(var="Site.Number")%>%
  setNames(c("Site.Number","upda_vep"))

all_list <- list(pecosclass_ca,pecosclass_upda,pecosclass_vep,ca_upda,ca_vep,upda_vep)

all_disagreement_site <- all_list%>%
  reduce(left_join,by="Site.Number")%>%
  rowwise()%>%
  mutate(mean_disagreement = mean(c_across(pecosclass_ca:upda_vep)))%>%
  ungroup()

most_disagree_site <- filter(all_disagreement_site, mean_disagreement > 200)%>%
  select(c("Site.Number","mean_disagreement"))
```

```{r}
ca_multi <- ca%>%
  rownames_to_column(var = "Site.Number")%>%
  mutate(across(-Site.Number, ~ if_else(. > 0, 1, 0)))%>%
  mutate(total_phases_ca = rowSums(select(., -Site.Number)))
  
vep_multi <- vep %>%
  mutate(across(-Site.Number, ~ if_else(. > 0, 1, 0)))%>%
  mutate(total_phases_vep = rowSums(select(., -Site.Number)))

pecosclass_multi <- pecosclass%>%
  mutate(total_phases_pecosclass = rowSums(select(., -Site.Number)))

ca_multi$distinct_components_ca <- apply(ca_multi[, !names(ca_multi) %in% c("Site.Number","total_phases_ca")], 1, function(x) {
  r <- rle(x)
  sum(r$values == 1)
})

vep_multi$distinct_components_vep <- apply(vep_multi[, !names(vep_multi) %in% c("Site.Number","total_phases_vep")], 1, function(x) {
  r <- rle(x)
  sum(r$values == 1)
})

pecosclass_multi$distinct_components_pecosclass <- apply(pecosclass_multi[, !names(pecosclass_multi) %in% c("Site.Number","total_phases_pecosclass")], 1, function(x) {
  r <- rle(x)
  sum(r$values == 1)
})

component_details <- merge(ca_multi,vep_multi,by="Site.Number")%>%
  merge(.,pecosclass_multi,by="Site.Number")%>%
  select(c("Site.Number","total_phases_ca","total_phases_vep","total_phases_pecosclass","distinct_components_ca","distinct_components_vep","distinct_components_pecosclass"))

mean(component_details$distinct_components_ca)
mean(component_details$distinct_components_vep)

disagree_and_component <- merge(most_disagree_site,component_details,by="Site.Number")

mean(disagree_and_component$distinct_components_ca)
mean(disagree_and_component$distinct_components_vep)
```

```{r}
disagree_ceramics <- sort_ceramics(master_pottery,disagree_components)%>%
  mutate(diff = Prop-master_pottery_plot$Prop)

master_ceramics_prop <- sort_ceramics(master_pottery,master_pottery)
```

```{r}
png('./FIGURES/Figure_7.png',height=10000,width=15000,res=1200)
ggplot()+
  geom_col(data=disagree_ceramics,aes(Type,diff,col=Ware,fill=Ware))+
  scale_color_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
  scale_fill_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
  labs(title = "Difference in Proportion of Ceramic Types by Ware between \nAll Sites andSites with the Most Disagreement",
       x=NULL,
       y="Difference")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.text.y=element_text(size=13),
        axis.text.x = element_text(angle = 45,hjust = 1,size=15),
        axis.title.y=element_text(size=15),
        plot.title=element_text(size=18),
        plot.margin = margin(15,5,5,5),
        legend.title=element_text(size=15),
        legend.text=element_text(size=13),
        panel.spacing.x = unit(0.5, "cm"))
dev.off()
```

## Supplemental

#### 6.5.2 Run chi squared of bivariate annual summaries

```{r}
#| warning: false

col_names <- c("pecosclass_CA","pecosclass_UPDA","pecosclass_VEP","CA_UPDA", "CA_VEP", "UPDA_VEP")

chi_square <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

# t <- matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2)

for (i in 1:nrow(chi_square)) {
  ### CHI SQUARED
  #pecosclass vs. CA
  chi_square[i,c("pecosclass_CA")] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. UPDA
  chi_square[i,c("pecosclass_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. VEP
  chi_square[i,c("pecosclass_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #CA vs. UPDA
  chi_square[i,c("CA_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #CA vs. VEP
  chi_square[i,c("CA_VEP")] <-  chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #UPDA vs. VEP
  chi_square[i,c("UPDA_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("UPDA")],site_threshold-summary[i,"UPDA"]),nrow = 2),correct=F)$p.value
}

write.csv(chi_square, "./OUTPUT/chi_squred_p-values_bivariate.csv",row.names=F)

chi_square_bivariate <- read.csv("./OUTPUT/chi_squred_p-values_bivariate.csv",row.names=NULL)%>%
  mutate(Year=750:1280)%>%
  gather(method,p_value,-Year)

chi_square_bivariate$method <- factor(chi_square_bivariate$method,levels=c("pecosclass_CA","pecosclass_UPDA","pecosclass_VEP","CA_UPDA","CA_VEP","UPDA_VEP"),labels=c("Pecos Class & CA","Pecos Class & UPDA","Pecos Class & VEP","CA & UPDA","CA & VEP","UPDA & VEP"))

png('./FIGURES/Supplemental_Figure_4.png',height=14000,width=24000,res=1200)
ggplot()+
  geom_line(data=chi_square_bivariate,aes(Year,p_value,col=method),lwd=1)+
  facet_wrap(~method)+
  geom_hline(yintercept=0.05,linetype="dotted",col="grey30",lwd=0.75)+
  labs(x="Year",
       y="P-value",
       title="Difference in Total Number of Present/Absent Sites for each Pair of Methods ")+
  scale_color_manual(values=c("#8F5D5D","#5F797B","#988675","#B1967D","#E9A377","#BABF95"))+
  theme_minimal()+
  theme(legend.position = "none",
        plot.title=element_text(size=18),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        strip.text=element_text(size=15),
        axis.title=element_text(size=13),
        axis.text=element_text(size=13),
        plot.margin = margin(2,0,0,2),
        panel.spacing.x = unit(0.5, "cm"))
dev.off()
```

#### 6.5.3 Run chi square of annual summaries for fifty year bins

```{r}
summary <- merge(pecosclass_binto50_sum,vep_binto50_sum,by="Year")%>%
  merge(.,upda_binto50_sum,by="Year")%>%
  merge(.,ca_binto50_sum,by="Year")%>%
  setNames(c("Year","pecosclass","VEP","UPDA","CA"))

fifty_years <- seq(775,1275, by=50)

summary <- subset(summary,Year %in% fifty_years)

chisq_p <- matrix(nrow = length(fifty_years), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- fifty_years

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"],summary[i,c("VEP")],site_threshold-summary[i,c("VEP")],summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")]),nrow = 2),correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values_fiftyyearbins.csv")
```

#### 6.5.4 Run chi squared of bivariate comparisons for fifty year bins

```{r}
chi_square <- as.data.frame(matrix(nrow = length(fifty_years), 
                                      ncol = length(col_names), 
                     dimnames = list(fifty_years, col_names)))


for (i in 1:nrow(chi_square)) {
  ### CHI SQUARED
  #pecosclass vs. CA
  chi_square[i,c("pecosclass_CA")] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. UPDA
  chi_square[i,c("pecosclass_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. VEP
  chi_square[i,c("pecosclass_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #CA vs. UPDA
  chi_square[i,c("CA_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #CA vs. VEP
  chi_square[i,c("CA_VEP")] <-  chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #UPDA vs. VEP
  chi_square[i,c("UPDA_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("UPDA")],site_threshold-summary[i,"UPDA"]),nrow = 2),correct=F)$p.value
}

write.csv(chi_square, "./OUTPUT/chi_squred_p-values_bivariate__fiftyyearbins.csv",row.names=F)
```

#### 6.5.5 Run chi square of annual summaries for pecos periods

```{r}
pecosclass_bins_summary <- merge(ca_bintopecosclass_sum,vep_bintopecosclass_sum,by="Year")%>%
  merge(.,upda_bintopecosclass_sum,by="Year")%>%
  merge(.,pecosclass_bintopecosclass_sum,by="Year")%>%
  setNames(c("Year","CA","VEP","UPDA","pecosclass"))

pp_years <- c(825,1025,1225)

summary <- subset(summary,Year %in% pp_years)

chisq_p <- matrix(nrow = length(pp_years), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- pp_years

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"],summary[i,c("VEP")],site_threshold-summary[i,c("VEP")],summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")]),nrow = 2),correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values_pecosperiods.csv")
```

#### 6.5.6 Run chi squared of bivariate comparisons for pecos periods

```{r}
chi_square <- as.data.frame(matrix(nrow = length(pp_years), 
                                      ncol = length(col_names), 
                     dimnames = list(pp_years, col_names)))


for (i in 1:nrow(chi_square)) {
  ### CHI SQUARED
  #pecosclass vs. CA
  chi_square[i,c("pecosclass_CA")] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. UPDA
  chi_square[i,c("pecosclass_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. VEP
  chi_square[i,c("pecosclass_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #CA vs. UPDA
  chi_square[i,c("CA_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #CA vs. VEP
  chi_square[i,c("CA_VEP")] <-  chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #UPDA vs. VEP
  chi_square[i,c("UPDA_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("UPDA")],site_threshold-summary[i,"UPDA"]),nrow = 2),correct=F)$p.value
}

write.csv(chi_square, "./OUTPUT/chi_squred_p-values_bivariate_pecosperiods.csv",row.names=F)
```

#### 6.5.5 Re-bin to 50 year sections

```{r}
pecosclass_binto50 <- rebin(pecosclass_annual,fifty_ranges)
vep_binto50 <- rebin(vep_annual,fifty_ranges)
upda_binto50 <- rebin(upda_annual,fifty_ranges)
ca_binto50 <- rebin(ca_annual,fifty_ranges)

pecosclass_binto50_sum <- pecosclass_binto50%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

vep_binto50_sum <- vep_bintoca%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

upda_binto50_sum <- upda_binto50%>%
  select(c("Site","Year","bins"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(bins,na.rm=T))

ca_binto50_sum <- ca_binto50%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))
```

#### 6.6.1 cohen's kappa

```{r}
#| warning: false 

row_names <- 750:1280
col_names <- c("pecosclass_CA","pecosclass_UPDA","pecosclass_VEP","CA_UPDA", "CA_VEP", "UPDA_VEP")

cohens_summary <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(cohens_summary )) {
  #pecosclass vs. CA
  cohens_summary[i,c("pecosclass_CA")] <- cohen_kappa(pecosclass_t,ca_t,i,site_threshold)
  
  #pecosclass vs. UPDA
  cohens_summary[i,c("pecosclass_UPDA")] <- cohen_kappa(pecosclass_t,upda_t,i,site_threshold)
  
  #pecosclass vs. VEP
  cohens_summary[i,c("pecosclass_VEP")] <- cohen_kappa(pecosclass_t,vep_t,i,site_threshold)

  #CA vs. UPDA
  cohens_summary[i,c("CA_UPDA")] <- cohen_kappa(ca_t,upda_t,i,site_threshold)
  
  #CA vs. VEP
  cohens_summary[i,c("CA_VEP")] <- cohen_kappa(ca_t,vep_t,i,site_threshold)
  
  #UPDA vs. VEP
  cohens_summary[i,c("UPDA_VEP")] <- cohen_kappa(upda_t,vep_t,i,site_threshold)
}

write.csv(cohens_summary, "./OUTPUT/cohens_kappa.csv",row.names=F)
```

```{r}
chisq_p1 <- read.csv("./OUTPUT/chi_squared_p-values.csv")

cohens_summary <- read.csv("./OUTPUT/cohens_kappa.csv")%>%
  mutate(Year=750:1280)

cohens_summary1<-cohens_summary%>%
  mutate(pecosclass_average_cohen = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         CA_average_cohen = (pecosclass_CA+CA_UPDA+CA_VEP)/3,
         UPDA_average_cohen = (pecosclass_UPDA+CA_UPDA+UPDA_VEP)/3,
         VEP_average_cohen = (pecosclass_VEP+CA_VEP+UPDA_VEP)/3,
         Year=750:1280)

cohens_pecosclass1 <- cohens_summary1%>%
  select(c("Year","pecosclass_average_cohen","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,cohen,-Year)

cohens_ca1 <- cohens_summary1%>%
  select(c("Year","CA_average_cohen","pecosclass_CA","CA_UPDA","CA_VEP"))%>%
  gather(Method,cohen,-Year)

cohens_upda1 <- cohens_summary1%>%
  select(c("Year","UPDA_average_cohen","pecosclass_UPDA","CA_UPDA","UPDA_VEP"))%>%
  gather(Method,cohen,-Year)

cohens_vep1 <- cohens_summary1%>%
  select(c("Year","VEP_average_cohen","pecosclass_VEP","CA_VEP","UPDA_VEP"))%>%
  gather(Method,cohen,-Year)



### REFACTOR FOR CLEAN PLOTTING
cohens_pecosclass1$Method <- factor(cohens_pecosclass1$Method,levels=c("pecosclass_CA","pecosclass_UPDA","pecosclass_VEP","pecosclass_average_cohen"))

cohens_ca1$Method <- factor(cohens_ca1$Method,levels=c("pecosclass_CA","CA_UPDA","CA_VEP","CA_average_cohen"))

cohens_upda1$Method <- factor(cohens_upda1$Method,levels=c("pecosclass_UPDA","CA_UPDA","UPDA_VEP","UPDA_average_cohen"))

cohens_vep1$Method <- factor(cohens_vep1$Method,levels=c("pecosclass_VEP","CA_VEP","UPDA_VEP","VEP_average_cohen"))
###



p1<-ggplot(data=chisq_p1,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  labs(x=NULL,
       y="Chi-Square P-value",
       title="Average Agreement (Cohen's Kappa) by Method")+
  theme_minimal()+
  theme(#legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p2<-ggplot(data=cohens_pecosclass1 ,aes(Year,cohen))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#6C6C8140","#6C6C8140","#6C6C8140","#3d405b"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y="Cohen's Kappa\n(Pecos Class)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p3<-ggplot(data=cohens_ca1 ,aes(Year,cohen))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F40","#E07A5F40","#E07A5F40","#E07A5F"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y="Cohen's Kappa\n(CA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p4<-ggplot(data=cohens_upda1 ,aes(Year,cohen))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#BCD1C580","#BCD1C580","#81B29A"))+
  labs(x=NULL,
       y="Cohen's Kappa\n(UPDA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p5<-ggplot(data=cohens_vep1 ,aes(Year,cohen))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BFA37366","#BFA37366","#BFA37366","#BFA373"))+
  labs(x="Year",
       y="Cohen's Kappa\n(VEP)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))


png('./FIGURES/Figure_5_CohensKappa.png',height=16000,width=12000,res=1200)
ggarrange(p1,p2,p3,p4,p5, nrow=5,ncol=1,align="hv",heights=c(0.75,1,1,1,1))
dev.off()
```

```{r}
cohens_summary <- read.csv("./OUTPUT/cohens_kappa.csv")%>%
  mutate(Year=750:1280)%>%
  gather(method,disagree,-Year)

pecosclass_ca <-filter(cohens_summary,method=="pecosclass_CA")
pecosclass_upda <-filter(cohens_summary,method=="pecosclass_UPDA")
pecosclass_vep <-filter(cohens_summary,method=="pecosclass_VEP")
ca_upda <-filter(cohens_summary,method=="CA_UPDA")
ca_vep <-filter(cohens_summary,method=="CA_VEP")
upda_vep <-filter(cohens_summary,method=="UPDA_VEP")


kk_av <- mean(cohens_summary$disagree)

p1<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_hline(yintercept=kk_av,linetype="dotted",linewidth=0.85,col="grey30")+
  geom_line(data=pecosclass_ca,aes(Year,disagree),col="#8F5D5D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y="Cohen's Kappa (anything above dotted line\nis better than average)",
       title ="Pecos Classs & CA")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title.y = element_text(size=15),
        axis.text.y=element_text(size=13),
        axis.text.x=element_blank(),
        plot.margin = margin(10,10,10,5))

p2<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_hline(yintercept=kk_av,linetype="dotted",linewidth=0.85,col="grey30")+
  geom_line(data=pecosclass_upda,aes(Year,disagree),col="#5F797B",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & UPDA")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p3<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#BFA373",fill="#F2CC8FA4",alpha=0.5,linewidth = 1)+
  geom_hline(yintercept=kk_av,linetype="dotted",linewidth=0.85,col="grey30")+
  geom_line(data=pecosclass_vep,aes(Year,disagree),col="#7E7267",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & VEP")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p4<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
    geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_hline(yintercept=kk_av,linetype="dotted",linewidth=0.85,col="grey30")+
  geom_line(data=ca_upda,aes(Year,disagree),col="#B1967D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y="Cohen's Kappa (anything above dotted line\nis better than average)",
       title ="CA & UPDA")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))


p5<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#BFA373",fill="#F2CC8FA4",alpha=0.5,linewidth = 1)+
  geom_hline(yintercept=kk_av,linetype="dotted",linewidth=0.85,col="grey30")+
  geom_line(data=ca_vep,aes(Year,disagree),col="#D08F69",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="CA & VEP")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p6<-ggplot()+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#BFA373",fill="#F2CC8FA4",alpha=0.5,linewidth = 1)+
  geom_hline(yintercept=kk_av,linetype="dotted",linewidth=0.85,col="grey30")+
  geom_line(data=upda_vep,aes(Year,disagree),col="#A0AB87",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="UPDA & VEP")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

png('./FIGURES/Figure_6_CohensKappa.png',height=14000,width=24000,res=1200)
ggarrange(p1,p2,p3,p4,p5,p6, nrow=2,ncol=3,align="hv")
dev.off()
```

### 6.6 Ceramics of Disagreeable Sites

1.  Find the most disagreeable sites
2.  Assess their assemblages

```{r}
pecosclass_disagree <- filter(pecosclass_annual,Site %in% disagree_and_component$Site.Number)
ca_disagree <- filter(ca_annual,Site %in% disagree_and_component$Site.Number)
upda_disagree <- filter(upda_annual,Site %in% disagree_and_component$Site.Number)
vep_disagree <- filter(vep_annual,Site %in% disagree_and_component$Site.Number)

pecosclass_disagree_sum <- pecosclass_disagree%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

ca_disagree_sum <- ca_disagree%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

upda_disagree_sum <- upda_disagree%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))

vep_disagree_sum <- vep_disagree%>%
  select(c("Site","Year","Occupation"))%>%
  group_by(Year)%>%
  summarise(total_count=sum(Occupation,na.rm=T))


p1<-ggplot(data=pecosclass_disagree_sum,aes(Year,total_count))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#3d405b",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,20),breaks=seq(0,20,by=5))+
  labs(title="Site Count of Most Disagreeable Sites (n=17)",
       x=NULL,
       y="Site Count\nPecos Classification")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2),
        plot.title=element_text(size=18))

p2<-ggplot(data=ca_disagree_sum ,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#E07A5F",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,20),breaks=seq(0,20,by=5))+
  labs(x=NULL,
       y="Site Count\n(CA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p3<-ggplot(data=upda_disagree_sum ,aes(Year,total_count))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#81B29A",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,20),breaks=seq(0,20,by=5))+
  labs(x=NULL,
       y="Site Count\n(UPDA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p4<-ggplot(data=vep_disagree_sum ,aes(Year,total_count))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=20,col=NA,fill="grey75",alpha=0.20)+
  geom_line(col="#BFA373",lwd=1.5)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,20),breaks=seq(0,20,by=5))+
  labs(x="Year",
       y="Site Count\n(VEP)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

# p5<-ggplot()+
#   geom_col(data=master_ceramics_prop,aes(Type,Prop,col=Ware,fill=Ware))+
#   scale_color_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
#   scale_fill_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
#   scale_y_continuous(limits=c(0,0.75),breaks=seq(0,0.75,by=0.25))+
#   labs(title= "Proportion of Ceramic Types by Ware for All Sites (n=79)",
#        x=NULL,
#        y="Proportion")+
#   theme_minimal()+
#   theme(panel.grid.minor = element_blank(),
#         panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
#         plot.title=element_text(size=18),
#         axis.title.x=element_blank(),
#         axis.title.y=element_text(size=15),
#         axis.text.x=element_blank(),
#         axis.text.y=element_text(size=13),
#         plot.margin = margin(15,5,5,5),
#         legend.title=element_text(size=15),
#         legend.text=element_text(size=13))
```

#### 6.10.1 Divide by method

```{r}
ceramic_disagreements <- all_list%>%
  reduce(left_join,by="Site.Number")

pecosclass_ca_disagree <- subset(ceramic_disagreements,pecosclass_ca >= as.numeric(stats[1,c("q3")]))

pecosclass_upda_disagree <- subset(ceramic_disagreements,pecosclass_upda >= as.numeric(stats[2,c("q3")]))

pecosclass_vep_disagree <- subset(ceramic_disagreements,pecosclass_vep >= as.numeric(stats[3,c("q3")]))

ca_upda_disagree <- subset(ceramic_disagreements,ca_upda >= as.numeric(stats[4,c("q3")]))

ca_vep_disagree <- subset(ceramic_disagreements,ca_vep >= as.numeric(stats[5,c("q3")]))

upda_vep_disagree <- subset(ceramic_disagreements,upda_vep >= as.numeric(stats[6,c("q3")]))



pecos_ca_disagree_ceramics <- sort_ceramics(master_pottery,pecosclass_ca_disagree)%>%
  mutate(pecosclass_ca = Prop-master_pottery_plot$Prop)%>%
  select(c("Type","Ware","pecosclass_ca"))

pecos_upda_disagree_ceramics <- sort_ceramics(master_pottery,pecosclass_upda_disagree)%>%
  mutate(pecosclass_upda = Prop-master_pottery_plot$Prop)%>%
  select(c("Type","Ware","pecosclass_upda"))

pecos_vep_disagree_ceramics <- sort_ceramics(master_pottery,pecosclass_vep_disagree)%>%
  mutate(pecosclass_vep = Prop-master_pottery_plot$Prop)%>%
  select(c("Type","Ware","pecosclass_vep"))

ca_upda_disagree_ceramics <- sort_ceramics(master_pottery,ca_upda_disagree)%>%
  mutate(ca_upda = Prop-master_pottery_plot$Prop)%>%
  select(c("Type","Ware","ca_upda"))

ca_vep_disagree_ceramics <- sort_ceramics(master_pottery,ca_vep_disagree)%>%
  mutate(ca_vep = Prop-master_pottery_plot$Prop)%>%
  select(c("Type","Ware","ca_vep"))

upda_vep_disagree_ceramics <- sort_ceramics(master_pottery,upda_vep_disagree)%>%
  mutate(upda_vep = Prop-master_pottery_plot$Prop)%>%
  select(c("Type","Ware","upda_vep"))


all_list_ceramics <- list(pecos_ca_disagree_ceramics,pecos_upda_disagree_ceramics,pecos_vep_disagree_ceramics,ca_upda_disagree_ceramics,ca_vep_disagree_ceramics,upda_vep_disagree_ceramics)

all_disagreement_ceramics <- all_list_ceramics%>%
  reduce(left_join,by=c("Type","Ware"))%>%
  gather(method,value,-c("Type","Ware"))


all_disagreement_ceramics$method <- factor(all_disagreement_ceramics$method,levels=c("pecosclass_ca","pecosclass_upda","pecosclass_vep","ca_upda","ca_vep","upda_vep"),labels=c("Pecos Class & CA","Pecos Class & UPDA","Pecos Class & VEP","CA & UPDA","CA & VEP","UPDA & VEP"))
```

```{r}
pecos_ca_disagree_ceramics <- sort_ceramics(master_pottery,pecosclass_ca_disagree)%>%
  select(c("Type","Ware","Prop"))%>%
  setNames(c("Type","Ware","pecosclass_ca"))

pecos_upda_disagree_ceramics <- sort_ceramics(master_pottery,pecosclass_upda_disagree)%>%
  select(c("Type","Ware","Prop"))%>%
  setNames(c("Type","Ware","pecosclass_upda"))

pecos_vep_disagree_ceramics <- sort_ceramics(master_pottery,pecosclass_vep_disagree)%>%
  select(c("Type","Ware","Prop"))%>%
  setNames(c("Type","Ware","pecosclass_vep"))

ca_upda_disagree_ceramics <- sort_ceramics(master_pottery,ca_upda_disagree)%>%
  select(c("Type","Ware","Prop"))%>%
  setNames(c("Type","Ware","ca_upda"))

ca_vep_disagree_ceramics <- sort_ceramics(master_pottery,ca_vep_disagree)%>%
  select(c("Type","Ware","Prop"))%>%
  setNames(c("Type","Ware","ca_vep"))

upda_vep_disagree_ceramics <- sort_ceramics(master_pottery,upda_vep_disagree)%>%
  select(c("Type","Ware","Prop"))%>%
  setNames(c("Type","Ware","upda_vep"))


all_list_ceramics <- list(pecos_ca_disagree_ceramics,pecos_upda_disagree_ceramics,pecos_vep_disagree_ceramics,ca_upda_disagree_ceramics,ca_vep_disagree_ceramics,upda_vep_disagree_ceramics)

all_disagreement_ceramics <- all_list_ceramics%>%
  reduce(left_join,by=c("Type","Ware"))%>%
  gather(method,value,-c("Type","Ware"))

all_disagreement_ceramics$method <- factor(all_disagreement_ceramics$method,levels=c("pecosclass_ca","pecosclass_upda","pecosclass_vep","ca_upda","ca_vep","upda_vep"),labels=c("Pecos Class & CA","Pecos Class & UPDA","Pecos Class & VEP","CA & UPDA","CA & VEP","UPDA & VEP"))
```

#### 6.10.2 Figure 7

```{r}
p1<-ggplot()+
  geom_col(data=master_pottery_plot,aes(Type,Prop,col=Ware,fill=Ware))+
  scale_color_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
  scale_fill_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
  scale_y_continuous(limits=c(0,0.75),breaks=seq(0,0.75,by=0.25))+
  labs(title= "Proportion of Ceramic Types by Ware for All Sites (n=79)",
       x=NULL,
       y="Proportion")+
  theme_minimal()+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=18),
        axis.title.y=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(15,5,5,5),
        legend.title=element_text(size=15),
        legend.text=element_text(size=13))

p2<-ggplot()+
  geom_col(data=all_disagreement_ceramics,aes(Type,value,col=Ware,fill=Ware))+
  scale_color_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
  scale_fill_manual(values=c("#c2c5aa","#023047","#52796f","#d4a373"))+
  labs(title = "Differences among Outliers of Pairwise Comparisons",
       x=NULL,
       y="Difference in Proportion by Ware")+
  facet_wrap(~method)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        strip.text=element_text(size=15),
        axis.text.y=element_text(size=13),
        axis.text.x = element_text(angle = 45,hjust = 1,size=10),
        axis.title.y=element_text(size=15),
        plot.title=element_text(size=18),
        plot.margin = margin(15,0,0,2),
        panel.spacing.x = unit(0.5, "cm"))

png('./FIGURES/Figure_7.png',height=14000,width=28000,res=1200)
ggarrange(p1,p2,nrow=2,ncol=1,common.legend=TRUE,legend="bottom",heights=c(1,2))
dev.off()
```

### 6.11 All Ceramics

```{r}
pecos_ceramics <- summarize_ceramics(pecosclass_binto50)
ca_ceramics <- summarize_ceramics(ca_binto50)
upda_ceramics <- summarize_ceramics(upda_binto50)
vep_ceramics <- summarize_ceramics(vep_binto50)

write.csv(pecos_ceramics,"./OUTPUT/pecos_ceramics.csv")
write.csv(ca_ceramics,"./OUTPUT/ca_ceramics.csv")
write.csv(upda_ceramics,"./OUTPUT/upda_ceramics.csv")
write.csv(vep_ceramics,"./OUTPUT/vep_ceramics.csv")
```

### 6.12 Ceramics of Most Disagreeable Sites

```{r}
pc_vep_disagreesites <- unique(pecosclass_vep_disagree$Site.Number)
ca_vep_disagreesites <- unique(ca_vep_disagree$Site.Number)
upda_vep_disagreesites <- unique(upda_vep_disagree$Site.Number)

total <- c(pc_vep_disagreesites,ca_vep_disagreesites,upda_vep_disagreesites)

disagreeable_sites <- as.data.frame(table(unlist(total)))

disagree_components <- filter(component_details,Site.Number %in% disagreeable_sites$Var1)%>%
  filter(distinct_components_vep>1)

vep_binto50_disagree <- filter(vep_binto50,Site %in% disagree_components$Site.Number)

vep_disagreeable_ceramics_binto50 <- summarize_ceramics(vep_binto50_disagree)

vep_disagreeable_occupation <- filter(vep,Site.Number %in% disagree_components$Site.Number)%>%
  mutate(across(-Site.Number, ~ if_else(. > 0, 1, 0)))

write.csv(vep_disagreeable_occupation,"./OUTPUT/VEP_disagreeablesites_occupation.csv")

write.csv(vep_disagreeable_ceramics_binto50,"./OUTPUT/VEP_disagreeablesites_ceramics.csv")
```
