---
title: "Untitled"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/sfield3/OneDrive - University of Wyoming/FIELD LAB HOME/PROJECT/SERIATION_MULTIVERSE")
```

```{r}
#| warning: false
packages <-c('terra','sf','ggplot2','tidyterra','dplyr','tidyr','stringr','gifski','gganimate','ggforce','ggpubr','scales','lmtest','tseries','FedData','tidyverse','compiler','msm','snowfall','parallel','doParallel','ggridges')

for(p in packages) if(p %in% rownames(installed.packages()) == F) { install.packages(p) }
for(p in packages) suppressPackageStartupMessages(library(p,quietly=T,character.only=T))

theme_set(theme_bw())
```

## Find out if any positive non integers in VEP

```{r}

vep_names <- vep$Site.Number
vep_names<-gsub("5MV0","5MV",vep_names)

vep_full <- read.csv("C:/Users/sfield3/Documents/RESEARCH/DATA GENERAL/SENSITIVE SITE LOCATIONS/db_vepii_n.csv")%>%
  filter(COsitenum %in% vep_names)
```

## Making ceramic cutoff larger

```{r}
# if cut off on ceramics is different
# df_diag <- subset(master_pottery, select=c(Site.Number,Chapin.Gray,Moccasin.Gray,Mancos.Gray,Mancos.Corrugated,Dolores.Corrugated,Mesa.Verde.Corrugated,Chapin.B.W,Piedra.B.W,Cortez.B.W,Mancos.B.W,McElmo.B.W,Mesa.Verde.B.W,Abajo.R.O,Bluff.B.R,Deadmans.B.R))%>%
# column_to_rownames("Site.Number")
# 
# # apply a total count filter 
# df_diag <- df_diag%>%
#   mutate(total = rowSums(across(where(is.numeric)),na.rm=T))%>%
#   filter(total >=25)%>%
#   select(-c(total))
# 
# 
# df_diag_allprop <-prop.table(as.matrix(df_diag),1)*100
# 
# ca <- nsj_ca_seriation(df_diag_allprop)
# 
# names <- rownames(ca)
# 
# master_pottery <- filter(master_pottery,Site.Number %in% names)
# vep <- filter(vep,Site.Number %in% names)
# survey <- filter(survey,Site.Number %in% names)
# sites <- filter(sites,NewSiteNum %in% names)
```

## Examine Longevity and Multi-component stuff

```{r}
vep_annual_test <- filter(vep_annual,Year>899 & Year <1151)%>%
  filter(.,Occupation == 1)

length(unique(vep_annual_test$Site))

output_sites <- unique(vep_annual_test$Site)

filters <- append(output_sites,"Year")


vep_annual_output <- vep_t[,filters]


count_runs_ones <- function(x) {
  r <- rle(x)
  sum(r$values == 1)
}

output <- as.data.frame(matrix(NA,60,2))
for(i in 1:60){
  test <- vep_annual_output[,i]
  test1 <- test%>%
    setNames(c("site"))
  test_c <- count_runs_ones(test1$site)
  output[i,1]<-colnames(test)
  output[i,2]<-count_runs_ones(test1$site)
}

test_site <- filter(vep_annual,Site=="5MV03540")

```

#### 6.5.1 Compare without Re-binning - missing Survey

```{r}
#| warning: false
# run chi square of annual summaries after rebinning
summary <- merge(ca_sum,upda_sum,by="Year")%>%
  merge(.,vep_sum,by="Year")%>%
  setNames(c("Year","CA","UPDA","VEP"))

s_summary <- summary[, c("Year","CA","UPDA","VEP")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values_NOsurvey.csv")
```

#### 6.5.1 Compare without Re-binning - missing CA

```{r}
#| warning: false
# run chi square of annual summaries after rebinning
summary <- merge(survey_sum,upda_sum,by="Year")%>%
  merge(.,vep_sum,by="Year")%>%
  setNames(c("Year","survey","UPDA","VEP"))

s_summary <- summary[, c("Year","survey","UPDA","VEP")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values_NOca.csv")
```

#### 6.5.1 Compare without Re-binning - missing UPDA

```{r}
#| warning: false
# run chi square of annual summaries after rebinning
summary <- merge(survey_sum,ca_sum,by="Year")%>%
  merge(.,vep_sum,by="Year")%>%
  setNames(c("Year","survey","CA","VEP"))

s_summary <- summary[, c("Year","survey","CA","VEP")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values_NOupda.csv")
```

#### 6.5.1 Compare without Re-binning - missing VEP

```{r}
#| warning: false
# run chi square of annual summaries after rebinning
summary <- merge(survey_sum,ca_sum,by="Year")%>%
  merge(.,upda_sum,by="Year")%>%
  setNames(c("Year","survey","CA","UPDA"))

s_summary <- summary[, c("Year","survey","CA","UPDA")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values_NOvep.csv")
```

```{r}
ceramic_disagreements <- all_list%>%
  reduce(left_join,by="Site.Number")%>%
  rowwise()%>%
  mutate(mean_disagreement = mean(c_across(pecosclass_ca:upda_vep)))%>%
  ungroup()

type_order <- c("Chapin.Gray","Chapin.B.W","Abajo.R.O","Piedra.B.W","Bluff.B.R","Moccasin.Gray","Mancos.Gray","Mancos.Corrugated","Cortez.B.W","Deadmans.B.R","Mancos.B.W","Dolores.Corrugated","McElmo.B.W","Mesa.Verde.Corrugated","Mesa.Verde.B.W")


md_pecosclass_ca <- filter(ceramic_disagreements,pecosclass_ca>200)
md_pecosclass_upda <- filter(ceramic_disagreements,pecosclass_upda>200)
md_pecosclass_vep <- filter(ceramic_disagreements,pecosclass_vep>200)
md_ca_upda <- filter(ceramic_disagreements,ca_upda>200)
md_ca_vep <- filter(ceramic_disagreements,ca_vep>200)
md_upda_vep <- filter(ceramic_disagreements,upda_vep>200)


pecosclass_ca_plot <- filter(master_pottery,Site.Number %in% md_pecosclass_ca$Site.Number)%>%
  select(-c("Site.Number","Source","Indeterminate.Gray","Indeterminate.Corrugated", "Indeterminate.White","Indeterminate.B.W","Indeterminate.B.R","Other"))%>%
 summarise(across(where(is.numeric),sum))%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column(var="Type")%>%
  setNames(c("Type","Count"))%>%
  mutate(Prop=Count/(sum(Count)))

pecosclass_upda_plot <- filter(master_pottery,Site.Number %in% md_pecosclass_upda$Site.Number)%>%
  select(-c("Site.Number","Source","Indeterminate.Gray","Indeterminate.Corrugated", "Indeterminate.White","Indeterminate.B.W","Indeterminate.B.R","Other"))%>%
 summarise(across(where(is.numeric),sum))%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column(var="Type")%>%
  setNames(c("Type","Count"))%>%
  mutate(Prop=Count/(sum(Count)))

pecosclass_vep_plot <- filter(master_pottery,Site.Number %in% md_pecosclass_vep$Site.Number)%>%
  select(-c("Site.Number","Source","Indeterminate.Gray","Indeterminate.Corrugated", "Indeterminate.White","Indeterminate.B.W","Indeterminate.B.R","Other"))%>%
 summarise(across(where(is.numeric),sum))%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column(var="Type")%>%
  setNames(c("Type","Count"))%>%
  mutate(Prop=Count/(sum(Count)))

ca_upda_plot <- filter(master_pottery,Site.Number %in% md_ca_upda$Site.Number)%>%
  select(-c("Site.Number","Source","Indeterminate.Gray","Indeterminate.Corrugated", "Indeterminate.White","Indeterminate.B.W","Indeterminate.B.R","Other"))%>%
 summarise(across(where(is.numeric),sum))%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column(var="Type")%>%
  setNames(c("Type","Count"))%>%
  mutate(Prop=Count/(sum(Count)))

ca_vep_plot <- filter(master_pottery,Site.Number %in% md_ca_vep$Site.Number)%>%
  select(-c("Site.Number","Source","Indeterminate.Gray","Indeterminate.Corrugated", "Indeterminate.White","Indeterminate.B.W","Indeterminate.B.R","Other"))%>%
 summarise(across(where(is.numeric),sum))%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column(var="Type")%>%
  setNames(c("Type","Count"))%>%
  mutate(Prop=Count/(sum(Count)))

upda_vep_plot <- filter(master_pottery,Site.Number %in% md_upda_vep$Site.Number)%>%
  select(-c("Site.Number","Source","Indeterminate.Gray","Indeterminate.Corrugated", "Indeterminate.White","Indeterminate.B.W","Indeterminate.B.R","Other"))%>%
 summarise(across(where(is.numeric),sum))%>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column(var="Type")%>%
  setNames(c("Type","Count"))%>%
  mutate(Prop=Count/(sum(Count)))

pecosclass_ca_plot$Type <- factor(pecosclass_ca_plot$Type,levels=type_order)
pecosclass_upda_plot$Type <- factor(pecosclass_upda_plot$Type,levels=type_order)
pecosclass_vep_plot$Type <- factor(pecosclass_vep_plot$Type,levels=type_order)
ca_upda_plot$Type <- factor(ca_upda_plot$Type,levels=type_order)
ca_vep_plot$Type <- factor(ca_vep_plot$Type,levels=type_order)
upda_vep_plot$Type <- factor(upda_vep_plot$Type,levels=type_order)


ggplot()+
  geom_col(data=pecosclass_ca_plot,aes(Type,Prop))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits=c(0,0.5))

ggplot()+
  geom_col(data=pecosclass_upda_plot,aes(Type,Prop))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits=c(0,0.5))

ggplot()+
  geom_col(data=pecosclass_vep_plot,aes(Type,Prop))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits=c(0,0.5))

ggplot()+
  geom_col(data=ca_upda_plot,aes(Type,Prop))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits=c(0,0.5))

ggplot()+
  geom_col(data=ca_vep_plot,aes(Type,Prop))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits=c(0,0.5))

ggplot()+
  geom_col(data=upda_vep_plot,aes(Type,Prop))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits=c(0,0.5))
```

### COMPARE CHI SQUARE WITHOUT CERTAIN METHODS

```{r}
#| warning: false

c_nopecosclass <- read.csv("./OUTPUT/archive/chi_squared_p-values_NOsurvey.csv")%>%
  select(c("V1","V2"))%>%
  setNames(c("Year","no_pecosclass"))

c_noca <- read.csv("./OUTPUT/archive/chi_squared_p-values_NOca.csv")%>%
  select(c("V1","V2"))%>%
  setNames(c("Year","no_CA"))

c_noupda <- read.csv("./OUTPUT/archive/chi_squared_p-values_NOupda.csv")%>%
  select(c("V1","V2"))%>%
  setNames(c("Year","no_UPDA"))

c_novep <- read.csv("./OUTPUT/archive/chi_squared_p-values_NOvep.csv")%>%
  select(c("V1","V2"))%>%
  setNames(c("Year","no_VEP"))

c_all <- merge(c_nopecosclass,c_noca,by="Year")%>%
  merge(.,c_noupda,by="Year")%>%
  merge(.,c_novep,by="Year")%>%
  gather(group,count,-Year)

c_all$group <-factor(c_all$group,levels=c("no_pecosclass","no_CA","no_UPDA","no_VEP"),labels=c("Pecos Class","CA","UPDA","VEP"))

png('./FIGURES/Supplemental_Figure_2.png',height=9000,width=13000,res=1200)
ggplot()+
  geom_line(data=c_all,aes(Year,count,col=group),linewidth=1)+
  geom_hline(yintercept=0.05,linetype="dotted",size=1)+
  scale_color_manual(values=c("#264653","#2a9d8f","#f4a261","#e76f51"))+
  labs(x="Year",
       y="Chi Square P-value",
       color="Chi Squared without...")+
  theme_minimal()+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))
dev.off()
```

### Percent Error when Sherd Count Increases

```{r}

ca_ranges<- data.frame(ranges = c("0700-0799","0800-0879","0880-0919","0920-0979","0980-1019","1020-1059","1060-1099","1100-1179","1180-1249","1250-1280"), 
           period = 1:10)

pecosclass_ranges <- data.frame(ranges=c("0700-0899","0900-999","0900-1149","1050-1149","1150-1300"),period=c("P1","P1P2","P2","P2P3","P3"))

vep_ranges <-data.frame(ranges=c("0700-0724","0725-0799","0800-0839","0840-0879","0880-0919","0920-0979","0980-1019","1020-1059","1060-1099","1100-1139","1140-1179","1180-1224","1225-1259","1260-1280"),period=c("ph6","ph7","ph8","ph9","ph10","ph11","ph12","ph13","ph14","ph15","ph16","ph17","ph18","ph19"))

cybersw_ranges <- read.csv("./DATA/Ceramics/UPDA_Ceramic_TypesandDatesfromCyberSW.csv")%>%
  select(c("Type","BegDate","EndDate"))%>%
  setNames(c("Type","Begin","End"))

upda_ranges <-data.frame(ranges=c("0700-0799","0800-0874","0875-0924","0925-0974","0975-1049","1050-1174","1175-1200","1201-1250","1251-1280"),period=c("ph1","ph2","ph3","ph4","ph5","ph6","ph7","ph8","ph9"))

pecosclass_ranges_plot <- pecosclass_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))%>%
  mutate(phase = c("PI","PI/PII","PII","PII/PIII","PIII"))

pecosclass_ranges_plot1 <- pecosclass_ranges_plot[c(1,3,5),]
pecosclass_ranges_plot2 <- pecosclass_ranges_plot[c(2,4),]

ca_ranges_plot <- ca_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))%>%
  mutate(phase = c("ph1","ph2","ph3","ph4","ph5","ph6","ph7","ph8","ph9","ph10"))

upda_ranges_plot <- upda_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))

vep_ranges_plot <- vep_ranges%>%
  separate(ranges,into=c("start","end"),sep="-")%>%
  mutate(start=as.numeric(start),
         end=as.numeric(end))
```

#### Five or more sherds

```{r}
percent_error <- read.csv("./OUTPUT/percent_error.csv")%>%
  mutate(Year=750:1280)%>%
  gather(method,disagree,-Year)

pecosclass_ca <-filter(percent_error,method=="pecosclass_CA")
pecosclass_upda <-filter(percent_error,method=="pecosclass_UPDA")
pecosclass_vep <-filter(percent_error,method=="pecosclass_VEP")
ca_upda <-filter(percent_error,method=="CA_UPDA")
ca_vep <-filter(percent_error,method=="CA_VEP")
upda_vep <-filter(percent_error,method=="UPDA_VEP")


p1<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_ca,aes(Year,disagree),col="#8F5D5D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y="Disagreement",
       title ="Pecos Classs & CA\n(5 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title.y = element_text(size=15),
        axis.text.y=element_text(size=13),
        axis.text.x=element_blank(),
        plot.margin = margin(10,10,10,5))

p2<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_upda,aes(Year,disagree),col="#5F797B",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & UPDA\n(5 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p3<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_vep,aes(Year,disagree),col="#988675",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & VEP\n(5 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p4<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
    geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_upda,aes(Year,disagree),col="#B1967D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y="Disagreement",
       title ="CA & UPDA\n(5 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))


p5<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_vep,aes(Year,disagree),col="#E9A377",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="CA & VEP\n(5 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p6<-ggplot()+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=upda_vep,aes(Year,disagree),col="#BABF95",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="UPDA & VEP\n(5 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))
```

#### Ten or more sherds

```{r}
percent_error <- read.csv("./OUTPUT/10 or more/percent_error.csv")%>%
  mutate(Year=750:1280)%>%
  gather(method,disagree,-Year)

pecosclass_ca <-filter(percent_error,method=="Survey.CA")
pecosclass_upda <-filter(percent_error,method=="Survey.UPDA")
pecosclass_vep <-filter(percent_error,method=="Survey.VEP")
ca_upda <-filter(percent_error,method=="UPDA.CA")
ca_vep <-filter(percent_error,method=="VEP.CA")
upda_vep <-filter(percent_error,method=="VEP.UPDA")


p7<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_ca,aes(Year,disagree),col="#8F5D5D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y="Disagreement",
       title ="Pecos Classs & CA\n(10 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title.y = element_text(size=15),
        axis.text.y=element_text(size=13),
        axis.text.x=element_blank(),
        plot.margin = margin(10,10,10,5))

p8<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_upda,aes(Year,disagree),col="#5F797B",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & UPDA\n(10 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p9<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_vep,aes(Year,disagree),col="#988675",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & VEP\n(10 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p10<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
    geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_upda,aes(Year,disagree),col="#B1967D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y="Disagreement",
       title ="CA & UPDA\n(10 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))


p11<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_vep,aes(Year,disagree),col="#E9A377",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="CA & VEP\n(10 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p12<-ggplot()+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=upda_vep,aes(Year,disagree),col="#BABF95",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="UPDA & VEP\n(10 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))
```

#### Twenty-five or more sherds

```{r}
percent_error <- read.csv("./OUTPUT/25 or more/percent_error.csv")%>%
  mutate(Year=750:1280)%>%
  gather(method,disagree,-Year)

pecosclass_ca <-filter(percent_error,method=="Survey.CA")
pecosclass_upda <-filter(percent_error,method=="Survey.UPDA")
pecosclass_vep <-filter(percent_error,method=="Survey.VEP")
ca_upda <-filter(percent_error,method=="UPDA.CA")
ca_vep <-filter(percent_error,method=="VEP.CA")
upda_vep <-filter(percent_error,method=="VEP.UPDA")


p13<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_ca,aes(Year,disagree),col="#8F5D5D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y="Disagreement",
       title ="Pecos Classs & CA\n(25 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title.y = element_text(size=15),
        axis.text.y=element_text(size=13),
        axis.text.x=element_blank(),
        plot.margin = margin(10,10,10,5))

p14<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_upda,aes(Year,disagree),col="#5F797B",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & UPDA\n(25 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p15<-ggplot()+
  geom_rect(data=pecosclass_ranges_plot1,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#3D405B",fill="#9898A2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=pecosclass_vep,aes(Year,disagree),col="#988675",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x=NULL,
       y=NULL,
       title ="Pecos Class & VEP\n(25 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p16<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
    geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_upda,aes(Year,disagree),col="#B1967D",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y="Disagreement",
       title ="CA & UPDA\n(25 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))


p17<-ggplot()+
  geom_rect(data=ca_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#E07A5F",fill="#E9B5A4",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=ca_vep,aes(Year,disagree),col="#E9A377",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="CA & VEP\n(25 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

p18<-ggplot()+
  geom_rect(data=upda_ranges_plot,aes(xmin=start,xmax=end,ymin=1.16,ymax=1.3),col="#81b29a",fill="#BAD1C2",alpha=0.5,linewidth = 1)+
  geom_rect(data=vep_ranges_plot,aes(xmin=start,xmax=end,ymin=1,ymax=1.14),col="#F2CC8F",fill="#F2DEBC",alpha=0.5,linewidth = 1)+
  geom_line(data=upda_vep,aes(Year,disagree),col="#BABF95",lwd=1.25)+
  scale_x_continuous(limits=c(675,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1.35),breaks=seq(0,1,by=0.5))+
  labs(x="Year",
       y=NULL,
       title ="UPDA & VEP\n(25 or more diagnostic sherds)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        plot.title=element_text(size=15),
        axis.title=element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(10,10,10,5))

```

```{r}
png('./FIGURES/Supplemental_Figure_3.png',height=14000,width=16000,res=600)
ggarrange(p1,p7,p13,
          p2,p8,p14,
          p3,p9,p15,
          p4,p10,p16,
          p5,p11,p17,
          p6,p12,p18, nrow=6,ncol=3,align="hv",heights=c(0.75,1,1,1,1,1))
dev.off()
```

### **Chi Squares for Disagreement**

```{r}
chi_square <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))
```

#### 6.9.1 Figure 8

```{r}
disagreement_no_mean <- filter(all_disagreement_site,method != "mean_disagreement")

disagreement_no_mean$method <- factor(disagreement_no_mean$method,levels=c("pecosclass_ca","pecosclass_upda","pecosclass_vep","ca_upda","ca_vep","upda_vep"),labels=c("Pecos Class & CA","Pecos Class & UPDA","Pecos Class & VEP","CA & UPDA","CA & VEP","UPDA & VEP"))

stats <- disagreement_no_mean%>%
  group_by(method)%>%
  summarise(mean= mean(value),
            q1=quantile(value,0.25),
            q3=quantile(value,0.75))


png('./FIGURES/Figure_8.png',height=8000,width=15000,res=1200)
ggplot()+
  geom_histogram(data=disagreement_no_mean,aes(x=value,color=method,fill=method),bins=15)+
  labs(x="Number of Years in Disagreement",
       y="Number of Sites")+
  scale_color_manual(values=c("#8F5D5D","#5F797B","#988675","#B1967D","#E9A377","#BABF95"))+
  scale_fill_manual(values=c("#8F5D5D","#5F797B","#988675","#B1967D","#E9A377","#BABF95"))+
  #scale_x_continuous(limits=c(0,300),breaks=seq(0,300,by=100))+
  geom_vline(data=stats,aes(xintercept=mean),linetype="dashed",lwd=0.8,col="grey10")+
  geom_vline(data=stats,aes(xintercept=q1),linetype="dotted",lwd=0.6,col="grey30")+
  geom_vline(data=stats,aes(xintercept=q3),linetype="dotted",lwd=0.6,col="grey30")+
  facet_wrap(~method)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        strip.text=element_text(size=15),
        axis.title=element_text(size=13),
        axis.text=element_text(size=13),
        plot.margin = margin(2,0,0,2),
        panel.spacing.x = unit(0.5, "cm"))
dev.off()
```

#### 6.6.3 Run chi squared of disagreements

```{r}
#example of the math

# acutal year
t <- contingency_table(pecosclass_t, ca_t, 300)
chisq.test(t)

# made up year with perfect agreement
t2 <- matrix(c(9,1,1,68),nrow=2)
chisq.test(t2)

# made up year neds to be reoriented for chi square to show non-meaningful difference of most agreeable outcome
t3 <- matrix(c(9,1,68,1),nrow=2)
chisq.test(t3)





t2 <- matrix(c(t[1,1],t[2,1],t[2,2],t[1,2]),nrow=2)

t3 <- matrix(c(49,1,28,1),nrow=2)

chisq.test(t3)
```

```{r}
#| warning: false

chi_square <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(chi_square)) {
  
  #pecosclass vs. CA
  t <- contingency_table(pecosclass_t, ca_t, i) 
  chi_square[i,c("pecosclass_CA")] <- chisq.test(matrix(c(t[1,1],t[2,1],t[2,2],t[1,2]),nrow=2),correct=F)$p.value
  
  #pecosclass vs. UPDA
  t <- contingency_table(pecosclass_t, upda_t, i)
  chi_square[i,c("pecosclass_UPDA")] <- chisq.test(matrix(c(t[1,1],t[2,1],t[2,2],t[1,2]),nrow=2),correct=F)$p.value
  
  #pecosclass vs. VEP
  t <- contingency_table(pecosclass_t, vep_t, i)
  chi_square[i,c("pecosclass_VEP")] <- chisq.test(matrix(c(t[1,1],t[2,1],t[2,2],t[1,2]),nrow=2),correct=F)$p.value
  
  #CA vs. UPDA
  t <- contingency_table(ca_t, upda_t, i)
  chi_square[i,c("CA_UPDA")] <- chisq.test(matrix(c(t[1,1],t[2,1],t[2,2],t[1,2]),nrow=2),correct=F)$p.value
  
  #CA vs. VEP
  t <- contingency_table(ca_t,vep_t, i)
  chi_square[i,c("CA_VEP")] <-  chisq.test(matrix(c(t[1,1],t[2,1],t[2,2],t[1,2]),nrow=2),correct=F)$p.value
  
  #UPDA vs. VEP
  t <- contingency_table(upda_t,vep_t, i)
  chi_square[i,c("UPDA_VEP")] <- chisq.test(matrix(c(t[1,1],t[2,1],t[2,2],t[1,2]),nrow=2),correct=F)$p.value
}

write.csv(chi_square, "./OUTPUT/chi_squred_p-values_percent_disagreements.csv",row.names=F)

chi_square_percdis <- read.csv("./OUTPUT/chi_squred_p-values_percent_disagreements.csv",row.names=NULL)%>%
  mutate(Year=750:1280)%>%
  gather(method,p_value,-Year)

ggplot()+
  geom_line(data=chi_square_percdis,aes(Year,p_value,col=method))+
  facet_wrap(~method)
```

# Supllemental Plot Figure 1?

```{r}
test <- read.csv("./OUTPUT/percent_error.csv")%>%
  gather(method,disagreement,-X)%>%
  setNames(c("Year","Pairwise","disagreement"))

ggplot()+
  geom_line(data=test,aes(Year,disagreement,col=Pairwise))+
  facet_wrap(~Pairwise)

test2 <- read.csv("./OUTPUT/chi_square_bivariate.csv")%>%
  gather(method,p_value,-X)%>%
  setNames(c("Year","Pairwise","p_value"))

ggplot()+
  geom_line(data=test2,aes(Year,p_value,col=Pairwise))+
  facet_wrap(~Pairwise)
```

```{r}
#| warning: false 
chisq_p1 <- read.csv("./OUTPUT/chi_squared_p-values.csv")
percent_error <- read.csv("./OUTPUT/percent_error.csv")%>%
  setNames(c("Year","pecosclass_VEP","pecosclass_UPDA","pecosclass_CA","VEP_UPDA","VEP_CA","UPDA_CA"))

disagreement_summary1<-percent_error%>%
  mutate(CA_average_error = (pecosclass_CA+UPDA_CA+VEP_CA)/3,
         pecosclass_average_error = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         VEP_average_error = (VEP_CA+VEP_UPDA+pecosclass_VEP)/3,
         UPDA_average_error = (pecosclass_UPDA+VEP_UPDA+UPDA_CA)/3,
         Year=750:1280)

disagreement_pecosclass1 <- disagreement_summary1%>%
  select(c("Year","pecosclass_average_error","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

disagreement_ca1 <- disagreement_summary1%>%
  select(c("Year","CA_average_error","pecosclass_CA","UPDA_CA","VEP_CA"))%>%
  gather(Method,error,-Year)

disagreement_upda1 <- disagreement_summary1%>%
  select(c("Year","UPDA_average_error","pecosclass_UPDA","VEP_UPDA","UPDA_CA"))%>%
  gather(Method,error,-Year)

disagreement_vep1 <- disagreement_summary1%>%
  select(c("Year","VEP_average_error","VEP_CA","VEP_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

p1<-ggplot(data=chisq_p1,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  labs(x=NULL,
       y="Chi-Square P-value",
       title="Average Disagreement by Method")+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p2<-ggplot(data=disagreement_pecosclass1 ,aes(Year,error))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#3d405b","#6C6C8140","#6C6C8140","#6C6C8140"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y="Disagreement\n(Pecos Class)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p3<-ggplot(data=disagreement_ca1 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F","#E07A5F40","#E07A5F40","#E07A5F40"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y="Disagreement\n(CA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p4<-ggplot(data=disagreement_upda1 ,aes(Year,error))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#81B29A","#BCD1C580","#BCD1C580"))+
  labs(x=NULL,
       y="Disagreement\n(UPDA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p5<-ggplot(data=disagreement_vep1 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#F5DEC0BF","#f2cc8f","#F5DEC0BF","#F5DEC0BF"))+
  labs(x="Year",
       y="Disagreement\n(VEP)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))


png('./FIGURES/Figure_6.png',height=16000,width=12000,res=1200)
ggarrange(p1,p2,p3,p4,p5, nrow=5,ncol=1,align="hv",heights=c(0.75,1,1,1,1))
dev.off()

table_4 <- t(sapply(disagreement_summary1, function(x) {
  c(mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE))
}))%>%
  as.data.frame()

write.csv(table_4,"./OUTPUT/error_nonbinned.csv")
```

#### 6.5.2 Compare Re-binned to pecosclass phases

```{r}
#| warning: false

# run chi square of annual summaries after rebinning
summary <- merge(ca_bintopecosclass_sum,vep_bintopecosclass_sum,by="Year")%>%
  merge(.,upda_bintopecosclass_sum,by="Year")%>%
  merge(.,pecosclass_bintopecosclass_sum,by="Year")%>%
  setNames(c("Year","CA","VEP","UPDA","pecosclass"))

s_summary <- summary[, c("Year","CA","pecosclass","VEP","UPDA")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}

### CHI.SQUARE BEING WEIGHTED BY VEP?
# mat <-matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
#                                        s_summary[i,3],site_threshold-s_summary[i,3], 
#                                        s_summary[i,4],site_threshold-s_summary[i,4], 
#                                        s_summary[i,5],site_threshold-s_summary[i,5]),
#                                      nrow = 2)
# 
# test <- mat[,c(1,3,4)]
# 
# chisq.test(test)

chisq_p2 <- as.data.frame(chisq_p)

write.csv(chisq_p2, "./OUTPUT/chi_squared_p-values_pecosclass_bins.csv")

# transform into site-by-site matrix
ca_t <- ca_bintopecosclass%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

vep_t <- vep_bintopecosclass%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

upda_t<- upda_bintopecosclass%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

pecosclass_t <- pecosclass_bintopecosclass%>%
  pivot_wider(names_from=Site,values_from=Occupation)

# calculate pairwise disagreement
#### Calculate pairwise disagreement ####
col_names <- c("pecosclass/VEP", "pecosclass/UPDA", "pecosclass/CA", "VEP/UPDA", "VEP/CA", "UPDA/CA")

percent_error <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(percent_error)) {
  #pecosclass vs. VEP
  percent_error[i,1] <- (contingency_table(pecosclass_t, vep_t, i)[1,2] + 
                           contingency_table(pecosclass_t, vep_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. UPDA
  percent_error[i,2] <- (contingency_table(pecosclass_t, upda_t, i)[1,2] + 
                           contingency_table(pecosclass_t, upda_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. CA
  percent_error[i,3] <- (contingency_table(pecosclass_t, ca_t, i)[1,2] + 
                           contingency_table(pecosclass_t, ca_t, i)[2,1]) /site_threshold
  
  #VEP vs. UPDA
  percent_error[i,4] <- (contingency_table(vep_t, upda_t, i)[1,2] + 
                           contingency_table(vep_t, upda_t, i)[2,1]) /site_threshold
  
  #VEP vs. CA
  percent_error[i,5] <- (contingency_table(vep_t, ca_t, i)[1,2] + 
                           contingency_table(vep_t, ca_t, i)[2,1]) /site_threshold
  
  #UPDA vs. CA
  percent_error[i,6] <- (contingency_table(upda_t, ca_t, i)[1,2] + 
                           contingency_table(upda_t, ca_t, i)[2,1]) /site_threshold
}

# setwd(outdir)
write.csv(percent_error, "./OUTPUT/percent_error_pecosclass_bins.csv")
```

#### 6.5.3 Compare Re-binned to CA phases

```{r}
#| warning: false

# run chi square of annual summaries after rebinning

summary <- merge(pecosclass_bintoca_sum,vep_bintoca_sum,by="Year")%>%
  merge(.,upda_bintoca_sum,by="Year")%>%
  merge(.,ca_bintoca_sum,by="Year")%>%
  setNames(c("Year","pecosclass","VEP","UPDA","CA"))

s_summary <- summary[, c("Year","CA","pecosclass","VEP","UPDA")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}


chisq_p3 <- as.data.frame(chisq_p)

write.csv(chisq_p3, "./OUTPUT/chi_squared_p-values_CA_bins.csv")

# transform into site-by-site matrix
pecosclass_t <- pecosclass_bintoca%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

vep_t <- vep_bintoca%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

upda_t<- upda_bintoca%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

ca_t <- ca_bintoca%>%
  pivot_wider(names_from=Site,values_from=Occupation)

# calculate pairwise diseagreement
#### Calculate pairwise disagreement ####
col_names <- c("pecosclass/VEP", "pecosclass/UPDA", "pecosclass/CA", "VEP/UPDA", "VEP/CA", "UPDA/CA")

percent_error <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(percent_error)) {
  #pecosclass vs. VEP
  percent_error[i,1] <- (contingency_table(pecosclass_t, vep_t, i)[1,2] + 
                           contingency_table(pecosclass_t, vep_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. UPDA
  percent_error[i,2] <- (contingency_table(pecosclass_t, upda_t, i)[1,2] + 
                           contingency_table(pecosclass_t, upda_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. CA
  percent_error[i,3] <- (contingency_table(pecosclass_t, ca_t, i)[1,2] + 
                           contingency_table(pecosclass_t, ca_t, i)[2,1]) /site_threshold
  
  #VEP vs. UPDA
  percent_error[i,4] <- (contingency_table(vep_t, upda_t, i)[1,2] + 
                           contingency_table(vep_t, upda_t, i)[2,1]) /site_threshold
  
  #VEP vs. CA
  percent_error[i,5] <- (contingency_table(vep_t, ca_t, i)[1,2] + 
                           contingency_table(vep_t, ca_t, i)[2,1]) /site_threshold
  
  #UPDA vs. CA
  percent_error[i,6] <- (contingency_table(upda_t, ca_t, i)[1,2] + 
                           contingency_table(upda_t, ca_t, i)[2,1]) /site_threshold
}

# setwd(outdir)
write.csv(percent_error, "./OUTPUT/percent_error_CA_bins.csv")
```

#### 6.5.4 Compare Re-binned to UPDA phases

```{r}
#| warning: false
# run chi square of annual summaries after rebinning
summary <- merge(pecosclass_bintoupda_sum,vep_bintoupda_sum,by="Year")%>%
  merge(.,ca_bintoupda_sum,by="Year")%>%
  merge(.,upda_bintoupda_sum,by="Year")%>%
  setNames(c("Year","pecosclass","VEP","UPDA","CA"))

s_summary <- summary[, c("Year","CA","pecosclass","VEP","UPDA")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}


chisq_p4 <- as.data.frame(chisq_p)

write.csv(chisq_p4, "./OUTPUT/chi_squared_p-values_UPDA_bins.csv")


# transform into site-by-site matrix
pecosclass_t <- pecosclass_bintoupda%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

vep_t<-vep_bintoupda%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

upda_t<- upda_bintoupda%>%
  pivot_wider(names_from=Site,values_from=Occupation)

ca_t <- ca_bintoupda%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

# calculate pairwise diseagreement
#### Calculate pairwise disagreement ####
col_names <- c("pecosclass/VEP", "pecosclass/UPDA", "pecosclass/CA", "VEP/UPDA", "VEP/CA", "UPDA/CA")

percent_error <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(percent_error)) {
  #pecosclass vs. VEP
  percent_error[i,1] <- (contingency_table(pecosclass_t, vep_t, i)[1,2] + 
                           contingency_table(pecosclass_t, vep_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. UPDA
  percent_error[i,2] <- (contingency_table(pecosclass_t, upda_t, i)[1,2] + 
                           contingency_table(pecosclass_t, upda_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. CA
  percent_error[i,3] <- (contingency_table(pecosclass_t, ca_t, i)[1,2] + 
                           contingency_table(pecosclass_t, ca_t, i)[2,1]) /site_threshold
  
  #VEP vs. UPDA
  percent_error[i,4] <- (contingency_table(vep_t, upda_t, i)[1,2] + 
                           contingency_table(vep_t, upda_t, i)[2,1]) /site_threshold
  
  #VEP vs. CA
  percent_error[i,5] <- (contingency_table(vep_t, ca_t, i)[1,2] + 
                           contingency_table(vep_t, ca_t, i)[2,1]) /site_threshold
  
  #UPDA vs. CA
  percent_error[i,6] <- (contingency_table(upda_t, ca_t, i)[1,2] + 
                           contingency_table(upda_t, ca_t, i)[2,1]) /site_threshold
}

# setwd(outdir)
write.csv(percent_error, "./OUTPUT/percent_error_UPDA_bins.csv")
```

#### 6.5.5 Compare Re-binned to VEP phases

```{r}
#| warning: false
# run chi square of annual summaries after rebinning

summary <- merge(ca_bintovep_sum,pecosclass_bintovep_sum,by="Year")%>%
  merge(.,upda_bintovep_sum,by="Year")%>%
  merge(.,vep_bintovep_sum,by="Year")%>%
  setNames(c("Year","CA","pecosclass","UPDA","VEP"))

s_summary <- summary[, c("Year","CA","pecosclass","VEP","UPDA")]

chisq_p <- matrix(nrow = nrow(s_summary), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(s_summary[i,2],site_threshold-s_summary[i,2], 
                                       s_summary[i,3],site_threshold-s_summary[i,3], 
                                       s_summary[i,4],site_threshold-s_summary[i,4], 
                                       s_summary[i,5],site_threshold-s_summary[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}


chisq_p5 <- as.data.frame(chisq_p)

write.csv(chisq_p5, "./OUTPUT/chi_squared_p-values_VEP_bins.csv")


# transform into site-by-site matrix
pecosclass_t <- pecosclass_bintovep%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

upda_t<- upda_bintovep%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

ca_t <- ca_bintovep%>%
  select(c("Site","Year","bins"))%>%
  pivot_wider(names_from=Site,values_from=bins)

vep_t <- vep_bintovep%>%
  pivot_wider(names_from=Site,values_from=Occupation)

# calculate pairwise diseagreement
#### Calculate pairwise disagreement ####
col_names <- c("pecosclass/VEP", "pecosclass/UPDA", "pecosclass/CA", "VEP/UPDA", "VEP/CA", "UPDA/CA")

percent_error <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(percent_error)) {
  #pecosclass vs. VEP
  percent_error[i,1] <- (contingency_table(pecosclass_t, vep_t, i)[1,2] + 
                           contingency_table(pecosclass_t, vep_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. UPDA
  percent_error[i,2] <- (contingency_table(pecosclass_t, upda_t, i)[1,2] + 
                           contingency_table(pecosclass_t, upda_t, i)[2,1]) /site_threshold
  
  #pecosclass vs. CA
  percent_error[i,3] <- (contingency_table(pecosclass_t, ca_t, i)[1,2] + 
                           contingency_table(pecosclass_t, ca_t, i)[2,1]) /site_threshold
  
  #VEP vs. UPDA
  percent_error[i,4] <- (contingency_table(vep_t, upda_t, i)[1,2] + 
                           contingency_table(vep_t, upda_t, i)[2,1]) /site_threshold
  
  #VEP vs. CA
  percent_error[i,5] <- (contingency_table(vep_t, ca_t, i)[1,2] + 
                           contingency_table(vep_t, ca_t, i)[2,1]) /site_threshold
  
  #UPDA vs. CA
  percent_error[i,6] <- (contingency_table(upda_t, ca_t, i)[1,2] + 
                           contingency_table(upda_t, ca_t, i)[2,1]) /site_threshold
}

# setwd(outdir)
write.csv(percent_error, "./OUTPUT/percent_error_VEP_bins.csv")
```

### 6.6 Contingency Comparison Supplemental Plots

#### 6.6.1 Plot Re-binned to pecosclass

```{r}

chisq_p2 <- read.csv("./OUTPUT/chi_squared_p-values_pecosclass_bins.csv")

percent_error <- read.csv("./OUTPUT/percent_error_pecosclass_bins.csv")%>%
  setNames(c("Year","pecosclass_VEP","pecosclass_UPDA","pecosclass_CA","VEP_UPDA","VEP_CA","UPDA_CA"))

disagreement_summary2<-percent_error%>%
  mutate(CA_average_error = (pecosclass_CA+UPDA_CA+VEP_CA)/3,
         pecosclass_average_error = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         VEP_average_error = (VEP_CA+VEP_UPDA+pecosclass_VEP)/3,
         UPDA_average_error = (pecosclass_UPDA+VEP_UPDA+UPDA_CA)/3,
         Year=750:1280)

table_sup_2 <- t(sapply(disagreement_summary2, function(x) {
  c(mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE))
}))%>%
  as.data.frame()

write.csv(table_sup_2,"./OUTPUT/error_pecosclass_bins.csv")

disagreement_pecosclass2 <- disagreement_summary2%>%
  select(c("Year","pecosclass_average_error","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

disagreement_ca2 <- disagreement_summary2%>%
  select(c("Year","CA_average_error","pecosclass_CA","UPDA_CA","VEP_CA"))%>%
  gather(Method,error,-Year)

disagreement_upda2 <- disagreement_summary2%>%
  select(c("Year","UPDA_average_error","pecosclass_UPDA","VEP_UPDA","UPDA_CA"))%>%
  gather(Method,error,-Year)

disagreement_vep2 <- disagreement_summary2%>%
  select(c("Year","VEP_average_error","VEP_CA","VEP_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

p6<-ggplot(data=chisq_p2,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  labs(x=NULL,
       y="Chi-Squared P-value",
       title="Binned to Pecos Class")+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p7<-ggplot(data=disagreement_pecosclass2 ,aes(Year,error))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#6C6C8140","#6C6C8140","#6C6C8140","#3d405b"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y="Disagreement\n(Pecos Class)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p8<-ggplot(data=disagreement_ca2 ,aes(Year,error))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F","#E07A5F40","#E07A5F40","#E07A5F40"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y="Disagreement\n(CA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p9<-ggplot(data=disagreement_upda2 ,aes(Year,error))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#BCD1C580","#81B29A","#BCD1C580"))+
  labs(x=NULL,
       y="Disagreement\n(UPDA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p10<-ggplot(data=disagreement_vep2 ,aes(Year,error))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#F5DEC0BF","#F5DEC0BF","#F5DEC0BF","#f2cc8f"))+
  labs(x="Year",
       y="Disagreement\n(VEP)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))
```

#### 6.6.2 Plot Re-binned to CA

```{r}

chisq_p3 <- read.csv("./OUTPUT/chi_squared_p-values_CA_bins.csv")

percent_error <- read.csv("./OUTPUT/percent_error_CA_bins.csv")%>%
  setNames(c("Year","pecosclass_VEP","pecosclass_UPDA","pecosclass_CA","VEP_UPDA","VEP_CA","UPDA_CA"))

disagreement_summary3<-percent_error%>%
  mutate(CA_average_error = (pecosclass_CA+UPDA_CA+VEP_CA)/3,
         pecosclass_average_error = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         VEP_average_error = (VEP_CA+VEP_UPDA+pecosclass_VEP)/3,
         UPDA_average_error = (pecosclass_UPDA+VEP_UPDA+UPDA_CA)/3,
         Year=750:1280)

table_sup_3 <- t(sapply(disagreement_summary3, function(x) {
  c(mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE))
}))%>%
  as.data.frame()

write.csv(table_sup_3,"./OUTPUT/error_CA_bins.csv")

disagreement_pecosclass3 <- disagreement_summary3%>%
  select(c("Year","pecosclass_average_error","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

disagreement_ca3 <- disagreement_summary3%>%
  select(c("Year","CA_average_error","pecosclass_CA","UPDA_CA","VEP_CA"))%>%
  gather(Method,error,-Year)

disagreement_upda3 <- disagreement_summary3%>%
  select(c("Year","UPDA_average_error","pecosclass_UPDA","VEP_UPDA","UPDA_CA"))%>%
  gather(Method,error,-Year)

disagreement_vep3 <- disagreement_summary3%>%
  select(c("Year","VEP_average_error","VEP_CA","VEP_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

p11<-ggplot(data=chisq_p3,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y=NULL,
       title="Binned to CA")+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p12<-ggplot(data=disagreement_pecosclass3 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#6C6C8140","#6C6C8140","#6C6C8140","#3d405b"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p13<-ggplot(data=disagreement_ca3 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F","#E07A5F40","#E07A5F40","#E07A5F40"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p14<-ggplot(data=disagreement_upda3 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#BCD1C580","#81B29A","#BCD1C580"))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p15<-ggplot(data=disagreement_vep3 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#F5DEC0BF","#F5DEC0BF","#F5DEC0BF","#f2cc8f"))+
  labs(x="Year",
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))
```

#### 6.6.3 Plot Re-binned to UPDA

```{r}
chisq_p4 <- read.csv("./OUTPUT/chi_squared_p-values_UPDA_bins.csv")

percent_error <- read.csv("./OUTPUT/percent_error_UPDA_bins.csv")%>%
  setNames(c("Year","pecosclass_VEP","pecosclass_UPDA","pecosclass_CA","VEP_UPDA","VEP_CA","UPDA_CA"))

disagreement_summary4<-percent_error%>%
  mutate(CA_average_error = (pecosclass_CA+UPDA_CA+VEP_CA)/3,
         pecosclass_average_error = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         VEP_average_error = (VEP_CA+VEP_UPDA+pecosclass_VEP)/3,
         UPDA_average_error = (pecosclass_UPDA+VEP_UPDA+UPDA_CA)/3,
         Year=750:1280)

table_sup_4 <- t(sapply(disagreement_summary4, function(x) {
  c(mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE))
}))%>%
  as.data.frame()

write.csv(table_sup_4,"./OUTPUT/error_UPDA_bins.csv")


disagreement_pecosclass4 <- disagreement_summary4%>%
  select(c("Year","pecosclass_average_error","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

disagreement_ca4 <- disagreement_summary4%>%
  select(c("Year","CA_average_error","pecosclass_CA","UPDA_CA","VEP_CA"))%>%
  gather(Method,error,-Year)

disagreement_upda4 <- disagreement_summary4%>%
  select(c("Year","UPDA_average_error","pecosclass_UPDA","VEP_UPDA","UPDA_CA"))%>%
  gather(Method,error,-Year)

disagreement_vep4 <- disagreement_summary4%>%
  select(c("Year","VEP_average_error","VEP_CA","VEP_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

p16<-ggplot(data=chisq_p4,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y=NULL,
       title="Binned to UPDA Phases")+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p17<-ggplot(data=disagreement_pecosclass4 ,aes(Year,error))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#6C6C8140","#6C6C8140","#6C6C8140","#3d405b"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p18<-ggplot(data=disagreement_ca4 ,aes(Year,error))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F","#E07A5F40","#E07A5F40","#E07A5F40"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p19<-ggplot(data=disagreement_upda4 ,aes(Year,error))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#BCD1C580","#81B29A","#BCD1C580"))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p20<-ggplot(data=disagreement_vep4 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#F5DEC0BF","#F5DEC0BF","#F5DEC0BF","#f2cc8f"))+
  labs(x="Year",
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))
```

#### 6.6.4 Plot Re-binned to VEP

```{r}

chisq_p5 <- read.csv("./OUTPUT/chi_squared_p-values_VEP_bins.csv")

percent_error <- read.csv("./OUTPUT/percent_error_VEP_bins.csv")%>%
  setNames(c("Year","pecosclass_VEP","pecosclass_UPDA","pecosclass_CA","VEP_UPDA","VEP_CA","UPDA_CA"))

disagreement_summary5<-percent_error%>%
  mutate(CA_average_error = (pecosclass_CA+UPDA_CA+VEP_CA)/3,
         pecosclass_average_error = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         VEP_average_error = (VEP_CA+VEP_UPDA+pecosclass_VEP)/3,
         UPDA_average_error = (pecosclass_UPDA+VEP_UPDA+UPDA_CA)/3,
         Year=750:1280)

table_sup_5 <- t(sapply(disagreement_summary5, function(x) {
  c(mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE))
}))%>%
  as.data.frame()

write.csv(table_sup_5,"./OUTPUT/error_VEP_bins.csv")


disagreement_pecosclass5 <- disagreement_summary5%>%
  select(c("Year","pecosclass_average_error","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

disagreement_ca5 <- disagreement_summary5%>%
  select(c("Year","CA_average_error","pecosclass_CA","UPDA_CA","VEP_CA"))%>%
  gather(Method,error,-Year)

disagreement_upda5 <- disagreement_summary5%>%
  select(c("Year","UPDA_average_error","pecosclass_UPDA","VEP_UPDA","UPDA_CA"))%>%
  gather(Method,error,-Year)

disagreement_vep5 <- disagreement_summary5%>%
  select(c("Year","VEP_average_error","VEP_CA","VEP_UPDA","pecosclass_VEP"))%>%
  gather(Method,error,-Year)

p21<-ggplot(data=chisq_p5,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y=NULL,
       title="Binned to VEP Phases")+
  theme_minimal()+
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p22<-ggplot(data=disagreement_pecosclass5 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#6C6C8140","#6C6C8140","#6C6C8140","#3d405b"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p23<-ggplot(data=disagreement_ca5 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F","#E07A5F40","#E07A5F40","#E07A5F40"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p24<-ggplot(data=disagreement_upda5 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#BCD1C580","#81B29A","#BCD1C580"))+
  labs(x=NULL,
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p25<-ggplot(data=disagreement_vep5 ,aes(Year,error))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#F5DEC0BF","#F5DEC0BF","#F5DEC0BF","#f2cc8f"))+
  labs(x="Year",
       y=NULL)+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))


png('./FIGURES/Supplemental_Figure.png',height=8000,width=15000,res=600)
ggarrange(p6,p11,p16,p21,
          p7,p12,p17,p22,
          p8,p13,p18,p23,
          p9,p14,p19,p24,
          p10,p15,p20,p25,nrow=5,ncol=4,align="hv",heights=c(0.75,1,1,1,1))
dev.off()

#adjustcolor("#F5DEC0",alpha=1)
```

### Extras

```{r}
chisq_5 <- read.csv("./OUTPUT/5 or more/chi_squared_p-values_pecosclass_bins.csv")%>%
  select(c("V1","V2"))%>%
  setNames(c("Year","5_or_more"))
chisq_10 <- read.csv("./OUTPUT/10 or more/chi_squared_p-values_pecosclass_bins.csv")%>%
  select(c("V1","V2"))%>%
  setNames(c("Year","10_or_more"))
chisq_25 <- read.csv("./OUTPUT/25 or more/chi_squared_p-values_pecosclass_bins.csv")%>%
  select(c("V1","V2"))%>%
  setNames(c("Year","25_or_more"))

all <- merge(chisq_5,chisq_10,by="Year")%>%
  merge(.,chisq_25,by="Year")%>%
  gather(group,count,-Year)


ggplot()+
  geom_line(data=all,aes(Year,count,col=group))
```

## Appendix A: Misc. Code

```{r}
#rectangles for stuff


# png('./FIGURES/Figure_UPDAtest.png',height=12000,width=24000,res=1200)
# ggplot(data=upda_annual,aes(Year,Occupation))+
#   geom_line()+
#   facet_wrap(~Site)
# dev.off()


# #pecosclass bins
#   annotate("rect",xmin=900,xmax=999,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)
#   
#   
# #cabins
#   annotate("rect",xmin=700,xmax=799,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)
#   
#   
# #vep bins
#   annotate("rect",xmin=700,xmax=724,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=800,xmax=839,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=880,xmax=919,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)
# 
# #upda bins
#   annotate("rect",xmin=800,xmax=874,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=925,xmax=974,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)+
#   annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=80,col=NA,fill="grey75",alpha=0.20)


# df_summary_plot <- merge(ca_sum,pecosclass_sum,by="Year")%>%
#   merge(.,vep_sum,by="Year")%>%
#   merge(.,upda_sum,by="Year")%>%
#   setNames(c("Year","CA","pecosclass","VEP","UPDA"))%>%
#   gather(Method,Count,-Year)
# 
# p1<-ggplot(data=df_summary_plot,aes(Year,Count))+
#   geom_line(aes(col=Method),lwd=1.5)+
#   scale_color_manual(values=c("#e07a5f","#3d405b","#81b29a","#f2cc8f"))+
#   labs(x=NULL,
#        y="Site Count")+
#   theme_minimal()+
#   theme(legend.position="none",
#         panel.grid.minor = element_blank(),
#         panel.grid.major = element_line(linewidth = 0.2, color = "grey50"),
#         axis.title.y = element_text(size=10),
#         plot.margin = margin(5,5,0,2))


### OR 

# df1_long <- pecosclass_t %>%
#   pivot_longer(-Year, names_to = "site", values_to = "val1")
# 
# df2_long <- ca_t %>%
#   pivot_longer(-Year, names_to = "site", values_to = "val2")
# 
# result <- df1_long %>%
#   inner_join(df2_long, by = c("Year", "site")) %>%
#   mutate(diff = abs(val1 - val2)) %>%
#   group_by(Year) %>%
#   slice_max(diff, n = 1, with_ties = FALSE)
# 
# site_summary <- df1_long %>%
#   inner_join(df2_long, by = c("Year", "site")) %>%
#   mutate(diff = abs(val1 - val2)) %>%
#   group_by(site) %>%
#   summarise(total_diff = sum(diff), .groups = "drop") %>%
#   arrange(desc(total_diff))
```

```{r}
{r}
pecosclass_bins_summary <- merge(ca_bintopecosclass_sum,vep_bintopecosclass_sum,by="Year")%>%
  merge(.,upda_bintopecosclass_sum,by="Year")%>%
  merge(.,pecosclass_bintopecosclass_sum,by="Year")%>%
  setNames(c("Year","CA","VEP","UPDA","pecosclass"))

pp_years <- c(825,1025,1225)

summary <- subset(summary,Year %in% pp_years)

chisq_p <- matrix(nrow = length(pp_years), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- pp_years

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"],summary[i,c("VEP")],site_threshold-summary[i,c("VEP")],summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")]),nrow = 2),correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values_pecosperiods.csv")
```

```{r}
chi_square <- as.data.frame(matrix(nrow = length(pp_years), 
                                      ncol = length(col_names), 
                     dimnames = list(pp_years, col_names)))


for (i in 1:nrow(chi_square)) {
  ### CHI SQUARED
  #pecosclass vs. CA
  chi_square[i,c("pecosclass_CA")] <- chisq.test(matrix(c(summary[i,c("CA")],site_threshold-summary[i,c("CA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. UPDA
  chi_square[i,c("pecosclass_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #pecosclass vs. VEP
  chi_square[i,c("pecosclass_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("pecosclass")],site_threshold-summary[i,"pecosclass"]),nrow = 2),correct=F)$p.value
  
  #CA vs. UPDA
  chi_square[i,c("CA_UPDA")] <- chisq.test(matrix(c(summary[i,c("UPDA")],site_threshold-summary[i,c("UPDA")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #CA vs. VEP
  chi_square[i,c("CA_VEP")] <-  chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("CA")],site_threshold-summary[i,"CA"]),nrow = 2),correct=F)$p.value
  
  #UPDA vs. VEP
  chi_square[i,c("UPDA_VEP")] <- chisq.test(matrix(c(summary[i,c("VEP")],site_threshold-summary[i,c("VEP")], summary[i,c("UPDA")],site_threshold-summary[i,"UPDA"]),nrow = 2),correct=F)$p.value
}

write.csv(chi_square, "./OUTPUT/chi_squred_p-values_bivariate_pecosperiods.csv",row.names=F)
```

```{r}
chisq_p1 <- read.csv("./OUTPUT/chi_squared_p-values.csv")

cohens_summary <- read.csv("./OUTPUT/cohens_kappa.csv")%>%
  mutate(Year=750:1280)

cohens_summary1<-cohens_summary%>%
  mutate(pecosclass_average_cohen = (pecosclass_CA+pecosclass_UPDA+pecosclass_VEP)/3,
         CA_average_cohen = (pecosclass_CA+CA_UPDA+CA_VEP)/3,
         UPDA_average_cohen = (pecosclass_UPDA+CA_UPDA+UPDA_VEP)/3,
         VEP_average_cohen = (pecosclass_VEP+CA_VEP+UPDA_VEP)/3,
         Year=750:1280)

cohens_pecosclass1 <- cohens_summary1%>%
  select(c("Year","pecosclass_average_cohen","pecosclass_CA","pecosclass_UPDA","pecosclass_VEP"))%>%
  gather(Method,cohen,-Year)

cohens_ca1 <- cohens_summary1%>%
  select(c("Year","CA_average_cohen","pecosclass_CA","CA_UPDA","CA_VEP"))%>%
  gather(Method,cohen,-Year)

cohens_upda1 <- cohens_summary1%>%
  select(c("Year","UPDA_average_cohen","pecosclass_UPDA","CA_UPDA","UPDA_VEP"))%>%
  gather(Method,cohen,-Year)

cohens_vep1 <- cohens_summary1%>%
  select(c("Year","VEP_average_cohen","pecosclass_VEP","CA_VEP","UPDA_VEP"))%>%
  gather(Method,cohen,-Year)



### REFACTOR FOR CLEAN PLOTTING
cohens_pecosclass1$Method <- factor(cohens_pecosclass1$Method,levels=c("pecosclass_CA","pecosclass_UPDA","pecosclass_VEP","pecosclass_average_cohen"))

cohens_ca1$Method <- factor(cohens_ca1$Method,levels=c("pecosclass_CA","CA_UPDA","CA_VEP","CA_average_cohen"))

cohens_upda1$Method <- factor(cohens_upda1$Method,levels=c("pecosclass_UPDA","CA_UPDA","UPDA_VEP","UPDA_average_cohen"))

cohens_vep1$Method <- factor(cohens_vep1$Method,levels=c("pecosclass_VEP","CA_VEP","UPDA_VEP","VEP_average_cohen"))
###



p1<-ggplot(data=chisq_p1,aes(V1,V2))+
  geom_line(col="grey30",lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,0.15),breaks=seq(0,0.15,by=0.05))+
  labs(x=NULL,
       y="Chi-Square P-value",
       title="Average Agreement (Cohen's Kappa) by Method")+
  theme_minimal()+
  theme(#legend.position="none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.title=element_text(size=15),
        plot.margin = margin(5,5,0,2))

p2<-ggplot(data=cohens_pecosclass1 ,aes(Year,cohen))+
  annotate("rect",xmin=900,xmax=999,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1149,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_color_manual(values=c("#6C6C8140","#6C6C8140","#6C6C8140","#3d405b"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  labs(x=NULL,
       y="Cohen's Kappa\n(Pecos Class)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p3<-ggplot(data=cohens_ca1 ,aes(Year,cohen))+
  annotate("rect",xmin=700,xmax=799,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1180,xmax=1249,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#E07A5F40","#E07A5F40","#E07A5F40","#E07A5F"))+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  labs(x=NULL,
       y="Cohen's Kappa\n(CA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p4<-ggplot(data=cohens_upda1 ,aes(Year,cohen))+
  annotate("rect",xmin=800,xmax=874,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=925,xmax=974,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1050,xmax=1174,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1201,xmax=1250,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BCD1C580","#BCD1C580","#BCD1C580","#81B29A"))+
  labs(x=NULL,
       y="Cohen's Kappa\n(UPDA)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))

p5<-ggplot(data=cohens_vep1 ,aes(Year,cohen))+
  annotate("rect",xmin=700,xmax=724,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=800,xmax=839,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=880,xmax=919,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=980,xmax=1019,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1060,xmax=1099,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1140,xmax=1179,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  annotate("rect",xmin=1225,xmax=1259,ymin=0,ymax=1,col=NA,fill="grey75",alpha=0.20)+
  geom_line(aes(col=Method),lwd=1.5)+
  scale_x_continuous(limits=c(700,1300),breaks=seq(700,1300,by=100))+
  scale_y_continuous(limits=c(0,1),breaks=seq(0,1,by=0.25))+
  scale_color_manual(values=c("#BFA37366","#BFA37366","#BFA37366","#BFA373"))+
  labs(x="Year",
       y="Cohen's Kappa\n(VEP)")+
  theme_minimal()+
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.2, color = "grey75"),
        axis.title.y = element_text(size=15),
        axis.text=element_text(size=13),
        plot.margin = margin(5,5,0,2))


png('./FIGURES/Figure_5_CohensKappa.png',height=16000,width=12000,res=1200)
ggarrange(p1,p2,p3,p4,p5, nrow=5,ncol=1,align="hv",heights=c(0.75,1,1,1,1))
dev.off()
```
