---
title: "Glowacki et al. 2026"
format: html
editor: visual
execute: 
  echo: false
---

# Research compendium for Glowacki et al. (2026)

This R-based research compendium supports the analyses from Seriation Multiverse, which shows that.....

View the raw source code and output at: provide link

**When using the code included in this research compendium, please cite *all* of the following:**

> Glowacki, D. M....

## 1.0 Set Up Environment

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "C:/Users/sfield3/OneDrive - University of Wyoming/FIELD LAB HOME/PROJECT/SERIATION_MULTIVERSE")
```

```{r}

packages <-c('terra','sf','ggplot2','tidyterra','dplyr','tidyr','gifski','gganimate','ggforce','ggpubr','scales','lmtest','tseries','FedData','tidyverse','compiler','msm','snowfall','parallel','doParallel')

for(p in packages) if(p %in% rownames(installed.packages()) == F) { install.packages(p) }
for(p in packages) suppressPackageStartupMessages(library(p,quietly=T,character.only=T))

theme_set(theme_bw())
```

### 1.1 Unique Functions

`contingency_table`

`BR`

`updf`: no change?

`nsj_ca_seriation`

```{r}
### FUNCTION #1:
contingency_table <- function(data1, data2, row) {
  contingency <- matrix(c(0,0,0,0), nrow = 2,
                        dimnames = list(dim1 = c("present", "absent"),
                                        dim2 = c("present", "absent")))
  for (i in 2:ncol(data1)) {
    if (data1[row, i] == 1) {
      if (data2[row, i] == 1) {contingency[1,1] <- contingency[1,1] + 1}
      else {contingency[1,2] <- contingency[1,2] + 1}
    } else {
      if (data2[row, i] == 1) {contingency[2,1] <- contingency[2,1] + 1}
      else {contingency[2,2] <- contingency[2,2] + 1}
    }
  }
  return (contingency)
}

### FUNCTION #2:
updf <- function(site,cer.type,ct,start,end,chron,interval=5,cutoff=0.1,min.period=25) {
  
  ## Remove any type with a count of 0 from all vectors
  trim <- which(ct>0)
  cer.type <- as.vector(cer.type[trim])
  site <- as.vector(site[1])
  start <- start[trim]
  end <- end[trim]
  ct <- ct[trim]
  chron <- chron[trim]
  
  
  ## define function for rounding by base defined in "interval" argument and round start and end dates for each type
  mround <- function(x,base){base*round(x/base)}
  start <- mround(start,interval)
  end <- mround(end,interval)
  
  ## find minimal ceramic intervals based on rounded type date overlaps
  years <- unique(sort(c(start,end)))
  period <- do.call(rbind,foreach(i=1:(length(years)-1)) %dopar% rbind(c(years[i],years[i+1]))) 
  # Ensure that no period is shorter than the specificed min.period argument and combine periods to satisfy this requirement
  if (length(period)>2) {
    per.short <- foreach(i=1:(nrow(period)-1)) %dopar% (period[i+1,1]-period[i,1])
    comb.per <- which(per.short<min.period)
    if (length(comb.per)>0){ period <- period[-comb.per,]
    for (i in 1:(nrow(period)-1)) {period[i,2] <- period[i+1,1]}}
  }  
  
  ## Define lists of ceramic period lengths and site period lengths
  cer.per <- foreach(i=1:length(start)) %dopar% (seq((start[i]:end[i]))+(start[i]-1))
  per.len <- foreach(i=1:nrow(period)) %dopar% (seq(period[i,1]:period[i,2])+(period[i,1])-1)
  
  ## Create Uniform Distance dataframe via parallel function
  ucalc <- function(a,b) {
    out <- (length(intersect(a,b))-1)/(length(a)-1)
    if (out<0) {out <- 0}
    return(out)}
  
  udist <- foreach(a=cer.per,.combine='rbind') %:% foreach(b=per.len,.combine='c') %dopar% (ucalc(a,b))
  if (length(udist)==1) {udist <- as.matrix(udist)}
  colnames(udist) <- paste('AD',period[,1],'-',period[,2],sep='')
  
  ## Create dataframe of prior values and sum to calculate prior probabilities
  prior.tab <- udist*ct
  if (ncol(prior.tab)>1) {
    if (length(prior.tab[which(chron==1),])==ncol(prior.tab)) {prior <- prior.tab[which(chron==1),]/sum(prior.tab[which(chron==1),])} else
    {prior <- colSums(prior.tab[which(chron==1),])/sum(prior.tab[which(chron==1),])}} else {
      prior <- udist[1,]
    }
  prior[is.na(prior)] <- 0
  
  ## Create dataframe of count probabilities
  pij <- sweep(prior.tab,2,colSums(prior.tab),'/')
  pij[is.na(pij)] <- 0
  
  ## Create dataframe of probabilities based on uniform distribution
  uij <- sweep(udist,2,colSums(udist),'/')
  uij[is.na(uij)] <- 0
  
  ## Create dataframe of standard deviations of uniform distribution probabilities
  sd.t <- sqrt(sweep((uij*(1-uij)),2,colSums(ceiling(udist)),'/'))
  
  ## Create dataframe of conditionals and calculate conditional probabilities
  cij <- dnorm(pij,uij,sd.t)
  cij[which(uij-sd.t==1)] <- dnorm(1,1,1) # For intervals with only one ceramic type, set to standard normal
  is.na(cij) <- !is.finite(cij) 
  
  # calculate conditional probability and rescale from 0-1
  conditional <- apply(cij,2,mean,na.rm=T)
  conditional[is.na(conditional)] <- 0
  if (sum(conditional)>0) {conditional <- conditional/sum(conditional)}
  
  ## Calculate posterior proportions and remove any generated NA
  posterior <- foreach(i=1:length(prior),.combine='c') %dopar% ((prior[i]*conditional[i]))
  posterior <- posterior/sum(posterior)
  posterior[is.na(posterior)] <- 0
  
  ## Deal with edge cases where sum of posterior probabilities = 0 or conditional = prior
  if(sum(posterior)==0) {posterior <- prior}
  if (identical(conditional,prior)) {posterior <- prior}
  
  # calculated beginning (lwr) and ending (upr) dates based on the first and last period with posterior probability about the selected threshold
  lwr <- period[min(which(posterior>cutoff*max(posterior))),1]
  upr <- period[max(which(posterior>cutoff*max(posterior))),2]
  
    ## Create output list and return
  out.list <- list(site=site,prior=prior,posterior=posterior,conditional=conditional,period=period,samp.size=sum(ct),occ=cbind(lwr,upr))
  return(out.list)
}

### FUNCTION #3:
BR <- function(x) {
  rd <- dim(x)[1]
  results <- matrix(0,rd,rd)
  for (s1 in 1:rd) {
    for (s2 in 1:rd) {
      x1Temp <- as.numeric(x[s1, ])
      x2Temp <- as.numeric(x[s2, ])
      br.temp <- 0
      results[s1,s2] <- 200 - (sum(abs(x1Temp - x2Temp)))}}
  row.names(results) <- row.names(x)
  colnames(results) <- row.names(x)
return(results)}

### FUNCTION #4:
nsj_ca_seriation <- function(test_ceramics){
  cal_allprop <- read.csv("https://raw.githubusercontent.com/sfield2/NSJ-MV-CeramicSeriation/refs/heads/main/DATA/VEP_calibration_proportions.csv",header=T,fileEncoding = 'UTF-8-BOM')%>%
    as.matrix()
  
  #### Merge calibration and sample data
  test <-rbind(cal_allprop,test_ceramics)

  #### Run BR and extract relevant data
  test_br_out <- as.data.frame(BR(test))
  test_br_out <- test_br_out[11:nrow(test),1:10]

  ### normalize
  seriation_occupation<- as.data.frame(matrix(NA,nrow(test_ceramics),10))
  colnames(seriation_occupation) <- 1:10
  rownames(seriation_occupation) <- rownames(test_ceramics)
  
  for(i in 1:nrow(seriation_occupation)){
    for(j in 1:ncol(seriation_occupation)){
      seriation_occupation[i,j] <- (test_br_out[i,j]-rowMeans(test_br_out[i,]))/sd(test_br_out[i,])}
  }
  
  rm(cal_allprop)
return(seriation_occupation)
}
```

## 2.0 Import Study Area Extent, Sites, and Raw Ceramic Counts

```{r}

sites <- st_read("./DATA/Spatial/ParkMesa_sites.gpkg")

master_pottery <- read.csv("./DATA/Ceramics/PM_ceramics.csv")
```

```{r}

ca_ranges<- data.frame(ranges = c("0700-0799","0800-0879","0880-0919","0920-0979","0980-1019","1020-1059","1060-1099","1100-1179","1180-1249","1250-1280"), 
           period = 1:10)

survey_ranges <- data.frame(ranges=c("0700-0899","0900-999","0900-1149","1050-1149","1150-1300"),period=c("P1","P1P2","P2","P2P3","P3"))

vep_ranges <-data.frame(ranges=c("0700-0724","0725-0799","0800-0839","0840-0879","0880-0919","0920-0979","0980-1019","1020-1059","1060-1099","1100-1139","1140-1179","1180-1224","1225-1259","1260-1280"),period=c("ph6","ph7","ph8","ph9","ph10","ph11","ph12","ph13","ph14","ph15","ph16","ph17","ph18","ph19"))
```

## 3.0 Read in Pre-analyzed Seriation Results

`vep`:

`survey`:

```{r}

vep <- read.csv("./DATA/Occupations/vep_output.csv")

survey <- read.csv("./DATA/Occupations/survey_output.csv")
```

## 4.0 Correspondence Analysis

Transform diagnostic sherds into proportions then run correspondence analysis.

```{r}

df_diag <- subset(master_pottery, select=c(Site.Number,Chapin.Gray,Moccasin.Gray,Mancos.Gray,Mancos.Corrugated,Dolores.Corrugated,Mesa.Verde.Corrugated,Chapin.B.W,Piedra.B.W,Cortez.B.W,Mancos.B.W,McElmo.B.W,Mesa.Verde.B.W,Abajo.R.O,Bluff.B.R,Deadmans.B.R))%>%
column_to_rownames("Site.Number")

df_diag_allprop <-prop.table(as.matrix(df_diag),1)*100

ca <- nsj_ca_seriation(df_diag_allprop)
```

## 5.0 UPDA

Spread master_pottery into types and assign counts

```{r}

cybersw_dates <- read.csv("./DATA/Ceramics/UPDA_Ceramic_TypesandDatesfromCyberSW.csv")%>%
  select(c("Type","BegDate","EndDate"))%>%
  setNames(c("Type","Begin","End"))

diagnostic_list <- c("Abajo.R.O","Bluff.B.R","Chapin.B.W","Chapin.Gray","Cortez.B.W","Deadmans.B.R","Dolores.Corrugated","Mancos.B.W","Mancos.Corrugated","Mancos.Gray","McElmo.B.W","Mesa.Verde.B.W","Mesa.Verde.Corrugated","Moccasin.Gray","Piedra.B.W")

upda_ceramics <- master_pottery%>%
  select(-c(Source))%>%
  gather(Type, Count,-Site.Number)%>%
  left_join(cybersw_dates,by="Type")%>%
  mutate(Chronology=ifelse(.$Type%in% diagnostic_list,1,0))%>%
  filter(Type!="Other",
         Count > 0)
```

```{r}

site_list <- unique(upda_ceramics$Site.Number)

## Run analysis for all sites
output_list <- list()

for (i in 1:length(site_list)) {
  site_sub <- upda_ceramics[which(upda_ceramics$Site == site_list[i]),]
  updf_out <- updf(site = site_sub$Site.Number,
                   cer.type=site_sub$Type,
                   ct=site_sub$Count,
                   start=site_sub$Begin,
                   end=site_sub$End,
                   chron=site_sub$Chronology,
                   interval=25,
                   cutoff=0.1,
                   min.period=10)
  output_list[[i]] <- updf_out
}

## Output estimated occupation ("occ" variable) for each site, read into csv
upda <- as.data.frame(matrix(nrow = 79, ncol = 2, dimnames = list(site_list, c('Start', 'End'))))

for (i in 1:length(site_list)) {
  upda[i,1] <- output_list[[i]]$occ[1]
  upda[i,2] <- output_list[[i]]$occ[2]
}
```

## 6.0 Analysis

### 6.1 Clean up environment

```{r}
rm(list=setdiff(ls(),c("ca","survey","upda","vep","contingency_table","ca_ranges","survey_ranges","vep_ranges")))
```

### 6.2 Transform Occupation assignments into Annual Reconstructions

```{r}
ca_annual <- ca%>%
  rownames_to_column(var = "Site.Number")%>%
  gather(key="Timeperiod", value = "Value",`1`:`10`)%>%
  mutate(range = ca_ranges$ranges[match(Timeperiod,ca_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise() %>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))%>%
  mutate(Occupation = ifelse(Occupation>0,1,0))%>%
  filter(Year>749,
         Year<1281)

survey_annual_nontransition <- survey%>%
  select(-c("P1P2","P2P3"))%>%
  gather(key="Timeperiod", value = "Value",-Site.Number)%>%
  mutate(range = survey_ranges$ranges[match(Timeperiod,survey_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise() %>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))

survey_annual_transition <- survey%>%
  select(c("Site.Number","P1P2","P2P3"))%>%
  gather(key="Timeperiod", value = "Value",-Site.Number)%>%
  mutate(range = survey_ranges$ranges[match(Timeperiod,survey_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise() %>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))

survey_annual <- full_join(survey_annual_nontransition,survey_annual_transition, by = c("Site","Year"))%>%
  mutate(Occupation = rowSums(cbind(Occupation.x,Occupation.y),na.rm=T))%>%
  select(-c("Occupation.x","Occupation.y"))%>% 
  mutate(Occupation = replace_na(Occupation, 0))%>%
  filter(Year>749,
         Year<1281)

vep_annual <- vep%>%
  gather(key="Timeperiod", value = "Value",-Site.Number)%>%
  mutate(range = vep_ranges$ranges[match(Timeperiod,vep_ranges$period)],
         start=substr(range,1L,4L),
         end=substr(range,6L,9L))%>%
  rowwise()%>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$start,.$end,by=1)))%>%
  mutate(Occupation = ifelse(Occupation>0,1,0))%>%
  filter(Year>749,
         Year<1281)


upda_annual <- upda%>%
  mutate(Value=1)%>%
  rownames_to_column(var = "Site.Number")%>%
  rowwise()%>%
  do(data.frame(Site = .$Site.Number, Occupation=.$Value,Year=seq(.$Start,.$End,by=1)))%>%
  complete(Site,Year=750:1300,fill=list(Occupation=0))%>%
  filter(Year>749,
         Year<1281)


rm(survey_annual_nontransition,survey_annual_transition)
```

### 6.3 Contingency Table 

```{r}
survey_t <- survey_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)

vep_t <- vep_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)

ca_t <- ca_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)

upda_t <- upda_annual%>%
  pivot_wider(names_from=Site,values_from=Occupation)


#### Calculate pairwise disagreement ####
col_names <- c("Survey/VEP", "Survey/UPDA", "Survey/CA", "VEP/UPDA", "VEP/CA", "UPDA/CA")
row_names <- 750:1280

percent_error <- as.data.frame(matrix(nrow = length(row_names), 
                                      ncol = length(col_names), 
                     dimnames = list(row_names, col_names)))

for (i in 1:nrow(percent_error)) {
  #Survey vs. VEP
  percent_error[i,1] <- (contingency_table(survey_t, vep_t, i)[1,2] + 
                           contingency_table(survey_t, vep_t, i)[2,1]) / 79
  
  #Survey vs. UPDA
  percent_error[i,2] <- (contingency_table(survey_t, upda_t, i)[1,2] + 
                           contingency_table(survey_t, upda_t, i)[2,1]) / 79
  
  #Survey vs. CA
  percent_error[i,3] <- (contingency_table(survey_t, ca_t, i)[1,2] + 
                           contingency_table(survey_t, ca_t, i)[2,1]) / 79
  
  #VEP vs. UPDA
  percent_error[i,4] <- (contingency_table(vep_t, upda_t, i)[1,2] + 
                           contingency_table(vep_t, upda_t, i)[2,1]) / 79
  
  #VEP vs. CA
  percent_error[i,5] <- (contingency_table(vep_t, ca_t, i)[1,2] + 
                           contingency_table(vep_t, ca_t, i)[2,1]) / 79
  
  #UPDA vs. CA
  percent_error[i,6] <- (contingency_table(upda_t, ca_t, i)[1,2] + 
                           contingency_table(upda_t, ca_t, i)[2,1]) / 79
}

# setwd(outdir)
write.csv(percent_error, "./OUTPUT/percent_error.csv")

```

# Create code to bring this together.

```{r}
counts <- read.csv("./DATA/counts_annual.csv")
```

```{r}

chisq_p <- matrix(nrow = nrow(counts), ncol = 2)
chisq_p[1:nrow(chisq_p)] <- row_names

for (i in 1:nrow(chisq_p)) {
  chisq_p[i, 2] <- chisq.test(matrix(c(counts[i,2], 79-counts[i,2], 
                                       counts[i,3], 79-counts[i,3], 
                                       counts[i,4], 79-counts[i,4], 
                                       counts[i,5], 79-counts[i,5]),
                                     nrow = 2),
                              correct = F)$p.value
}

write.csv(chisq_p, "./OUTPUT/chi_squared_p-values.csv")

```
